---
title: "data_riskSLIM"
output: html_notebook
---

In this notebook we convert the features_before_on (stored in expanded_features.Rdata) to a format suitable for riskSLIM, and convert to two csvs: a training set and a held out test (same as the one used in predict_recidivism.Rmd)

riskSLIM load_data_from_csv() docstring: 
    """
    Parameters
    ----------
    dataset_csv_file                csv file containing the training data
                                    see /datasets/adult_data.csv for an example
                                    training data stored as a table with N+1 rows and d+1 columns
                                    column 1 is the outcome variable entries must be (-1,1) or (0,1)
                                    column 2 to d+1 are the d input variables
                                    row 1 contains unique names for the outcome variable, and the input variable

    sample_weights_csv_file         csv file containing sample weights for the training data
                                    weights stored as a table with N rows and 1 column
                                    all sample weights must be non-negative

    fold_csv_file                   csv file containing indices of folds for K-fold cross validation
                                    fold indices stored as a table with N rows and 1 column
                                    folds must be integers between 1 to K
                                    if fold_csv_file is None, then we do not use folds

    fold_num                        int between 0 to K, where K is set by the fold_csv_file
                                    let fold_idx be the N x 1 index vector listed in fold_csv_file
                                    samples where fold_idx == fold_num will be used to test
                                    samples where fold_idx != fold_num will be used to train the model
                                    fold_num = 0 means use "all" of the training data (since all values of fold_idx \in [1,K])
                                    if fold_csv_file is None, then fold_num is set to 0

    Returns
    -------
    dictionary containing training data for a binary classification problem with the fields:

     - 'X' N x P matrix of features (numpy.ndarray) with a column of 1s for the '(Intercept)'
     - 'Y' N x 1 vector of labels (+1/-1) (numpy.ndarray)
     - 'variable_names' list of strings containing the names of each feature (list)
     - 'Y_name' string containing the name of the output (optional)
     - 'sample_weights' N x 1 vector of sample weights, must all be positive

    """

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(caret)

```


```{r}
data_path = ""
load(paste0(data_path,"expanded_features.Rdata"))
load(paste0(data_path,"compas_psa.Rdata"))

```

Using compas and PSA features in addition to features we computed ourselves. 
```{r}
### Add useful columns to features and apply row filters used for all models
features_filt = features_before_on %>%
  inner_join(
    psa_features%>%
      select(-c(p_current_age,p_prison)), 
    by = c("person_id","screening_date"))%>%
  inner_join(outcomes, by = c("person_id","screening_date")) %>%
  filter(`Risk of Recidivism_decile_score` != -1, `Risk of Violence_decile_score` != -1) %>% # Filter 1
  filter(!is.na(current_offense_date)) %>% # Filter 3
  filter(screening_date <= current_offense_date_limit) %>% # Filter 4
  mutate(recid_use = recid) # Select recidivism or violent recidivism to use in this script 
         
## Select features and round count features
df = features_filt %>%
  transmute(
    #COMPAS Risk of Recidivism Features (excluding age)
    person_id,
    p_current_age,
    p_charge = pmin(p_charge, 5),
    p_jail30 = pmin(p_jail30, 5),
    p_prison = pmin(p_prison, 5),
    p_probation = pmin(p_probation, 5),
    
    #COMPAS Risk of violent recidivism features
    p_juv_fel_count = pmin(p_juv_fel_count, 5),
    p_felprop_violarrest = pmin(p_felprop_violarrest, 5),
    p_murder_arrest = pmin(p_murder_arrest, 5),
    p_felassault_arrest = pmin(p_felassault_arrest, 5),
    p_misdemassault_arrest = pmin(p_misdemassault_arrest, 5),
    p_famviol_arrest,
    p_sex_arrest,
    p_weapons_arrest,
    
      #PSA Features (which were not named above)
    fail_appear_two_yr, 
    fail_appear_two_plus,
    current_violent, 
    current_violent20, 
    pending_charge, 
    prior_conviction_F, 
    prior_conviction_M = pmin(prior_conviction_M, 5), 
    violent_conviction, 
    total_convictions = pmin(total_convictions, 5), 

    #Misc Features
    p_arrest,
    p_property = pmin(p_property, 5),
    p_traffic,
    p_drug,
    p_dui,
    p_domestic,
    p_stalking,
    p_voyeurism,
    p_fraud,
    p_stealing,
    p_trespass,
    
    #Binning age features 
    p_current_age18 = if_else(p_current_age==18, 1, 0),
    p_current_age1929 = if_else(p_current_age>18 && p_current_age<=29, 1, 0),
    p_current_age3039 = if_else(p_current_age>29 && p_current_age <=39, 1, 0),
    p_current_age4049 = if_else(p_current_age>39 && p_current_age <=49, 1, 0),
    p_current_age5059 = if_else(p_current_age>49 && p_current_age <=59, 1, 0),
    p_current_age60plus = if_else(p_current_age>=60, 1, 0),

    # p_current_age27 = if_else(p_current_age<=27, 1, 0),
    # p_current_age29 = if_else(p_current_age<=29, 1, 0),
    # 
    # p_current_age32 = if_else(p_current_age<=32, 1, 0),
    # p_current_age35 = if_else(p_current_age<=35, 1, 0),
    # p_current_age38 = if_else(p_current_age<=38, 1, 0),
    # 
    # p_current_age42 = if_else(p_current_age<=42, 1, 0),
    # p_current_age46 = if_else(p_current_age<=46, 1, 0),
    # p_current_age51 = if_else(p_current_age<=51, 1, 0),
    # p_current_age56 = if_else(p_current_age<=56, 1, 0),
    # p_current_age62 = if_else(p_current_age<=62, 1, 0),
    # p_current_age68 = if_else(p_current_age<=68, 1, 0),
    # p_current_age75 = if_else(p_current_age<=75, 1, 0),

    # p_age_first_offense21 = if_else(p_age_first_offense<=21, 1, 0),
    # p_age_first_offense23 = if_else(p_age_first_offense<=23, 1, 0),
    # p_age_first_offense25 = if_else(p_age_first_offense<=25, 1, 0),
    # p_age_first_offense27 = if_else(p_age_first_offense<=27, 1, 0),
    # p_age_first_offense29 = if_else(p_age_first_offense<=29, 1, 0),
    # 
    # p_age_first_offense32 = if_else(p_age_first_offense<=32, 1, 0),
    # p_age_first_offense35 = if_else(p_age_first_offense<=35, 1, 0),
    # p_age_first_offense38 = if_else(p_age_first_offense<=38, 1, 0),
    # 
    # p_age_first_offense42 = if_else(p_age_first_offense<=42, 1, 0),
    # p_age_first_offense46 = if_else(p_age_first_offense<=46, 1, 0),
    # p_age_first_offense51 = if_else(p_age_first_offense<=51, 1, 0),
    # p_age_first_offense56 = if_else(p_age_first_offense<=56, 1, 0),
    # p_age_first_offense62 = if_else(p_age_first_offense<=62, 1, 0),
    # p_age_first_offense68 = if_else(p_age_first_offense<=68, 1, 0),
    # p_age_first_offense75 = if_else(p_age_first_offense<=75, 1, 0),
    # 
    p_property0 = if_else(p_property == 0, 1, 0),
    p_property13 = if_else(p_property >= 1 && p_property <=3, 1, 0),
    p_property46 = if_else(p_property >= 4 && p_property <=6, 1, 0),
    p_property7up = if_else(p_property >= 7, 1, 0),

    prior_conviction_M0= if_else(prior_conviction_M == 0, 1, 0),
    prior_conviction_M13 = if_else(prior_conviction_M >= 1 && prior_conviction_M <=3, 1, 0),
    prior_conviction_M46 = if_else(prior_conviction_M >= 4 && prior_conviction_M <=6, 1, 0),
    prior_conviction_M7up = if_else(prior_conviction_M >= 7, 1, 0),
    
    p_charge0 = if_else(p_charge ==0, 1, 0),
    p_charge15 = if_else(p_charge >= 1 && p_charge <=5, 1, 0),
    p_charge610 = if_else(p_charge >= 6 && p_charge <=10, 1, 0),
    p_charge1120 = if_else(p_charge >=11  && p_charge <=20, 1, 0),
    p_charge21up = if_else(p_charge >=21, 1, 0),
    
    p_felprop_violarrest0= if_else(p_felprop_violarrest == 0, 1, 0),
    p_felprop_violarrest14 = if_else(prior_conviction_M >= 1 && prior_conviction_M <=4, 1, 0),
    p_felprop_violarrest5up = if_else(prior_conviction_M >= 5, 1, 0),
    
    total_convictions0= if_else(total_convictions == 0, 1, 0),
    total_convictions13 = if_else(total_convictions >= 1 && total_convictions <=3, 1, 0),
    total_convictions46 = if_else(total_convictions >=4  && total_convictions <=6, 1, 0),
    total_convictions7up = if_else(total_convictions >=7, 1, 0),

    #refactor labelfrom {0,1} to {-1,1} 
    recid_use = ifelse(recid_use == 0, -1, recid_use)) 
    # na.omit() #CREATES A LOT OF NAs

#label must be first column for riskSLIM
set.seed(283)
train = df %>% sample_frac(.8)
test = df %>% anti_join(train, by = 'person_id') #change to account for screening date!

train = select(train, -person_id)
test = select(test, -person_id)

train=train[,c(ncol(train),2:ncol(train)-1)]
test=test[,c(ncol(test),2:ncol(test)-1)]

```

Create K folds csv. riskSLIM pkg requires K \in [1,N], where N is the number of data points
```{r}
# folds=as.data.frame(createFolds(train$recid_use, k = 5, list = F, returnTrain = T),make.names = F)

```

```{r}
write.table(train, file="riskSLIM_train_data.csv",
            sep=",", row.names = F)
write.table(test, file="riskSLIM_test_data.csv",
            sep=",", row.names = F)

# write.table(folds,file= "riskSLIM_folds.csv",
#             sep = ",", col.names = F,row.names =F)

```

