---
title: "EDA"
author: "Beau Coker"
date: "1/14/2019"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
# Edit path_data to the directory with the kentucky data
# path_data = '/Volumes/William H Taft/KY Pretrial + Court Data FY 10-18/Data/'
path_data = '../KY Pretrial + Court Data FY 10-18/'
```

Note: all of the data files have a line like "\n (42352 row(s) affected) \n" at the end of the file. I manually removed this part to make it easier to read in. The code should run anyway but there will be a row of NAs in the dataframe if you don't remove this part

## Defendants

Includes Race, Ethnicity, Gender, and data source. Each row is a person 

```{r}
data_def = read_delim(paste0(path_data,'01_KY_Recidivism_Defendants.txt'), '|', col_names=TRUE) %>%
  mutate_if(is.character, str_trim)

data_def
```

No PersonID appears twice:
```{r}
length(unique(data_def$PersonID)) == nrow(data_def)
```


## 02 Interviews (int)

```{r}
data_int = read_delim(paste0(path_data,'02_KY_Recidivism_Interviews.txt'), '|', col_names=TRUE) %>%
  mutate_if(is.character, str_trim)

data_int
```

```{r}
names(data_int)
```

```{r}
data_int %>%
  mutate_all(is.null) %>%
  summarize_all(sum)
```



PersonIds often appear more than once:
```{r}
data_int %>%
  group_by(PersonId) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

### Compute Arnold PSA


```{r}
# Add Felony or Misdemeanor column
data_int = data_int %>%
  mutate(RA2_Has_Prior_Felony_Misdemeanor = if_else(RA2_Has_Prior_Misdemeanor=="NULL" | RA2_Has_Prior_Felony=="NULL","NULL",
                                                    if_else(RA2_Has_Prior_Misdemeanor=="2" | RA2_Has_Prior_Felony=="2","2","1")))
```

```{r}
# Failure to Appear (FTA)
FTA = data_int %>%
  transmute(
    RA2_New_Charge_While_Case_Pending = factor(RA2_New_Charge_While_Case_Pending, levels=c("1","2"), labels=c(1,0)),
    RA2_Has_Prior_Felony_Misdemeanor = factor(RA2_Has_Prior_Felony_Misdemeanor, levels=c("1","2"), labels=c(1,0)),
    RA2_How_Many_FTA_Last_2_Yrs = factor(RA2_How_Many_FTA_Last_2_Yrs, levels=c("1","2","3"), labels=c(0,1,2)), # Questions don't match documentation exactly
    RA2_Has_Any_FTA_Older_2_Yrs = factor(RA2_Has_Any_FTA_Older_2_Yrs, levels=c("1","2"), labels=c(1,0))
  ) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

FTA$score = apply(FTA,1,sum)
```


```{r}
# New Criminal Activity (NCA)
NCA = data_int %>%
  transmute(
    RA2_Age_At_Interview = factor(RA2_Age_At_Interview, levels=c("1","2","3"), labels=c(2,2,0)),
    RA2_New_Charge_While_Case_Pending = factor(RA2_New_Charge_While_Case_Pending, levels=c("1","2"), labels=c(3,0)),
    RA2_Has_Prior_Misdemeanor = factor(RA2_Has_Prior_Misdemeanor, levels=c("1","2"), labels=c(1,0)),
    RA2_Has_Prior_Felony = factor(RA2_Has_Prior_Felony, levels=c("1","2"), labels=c(1,0)),
    RA2_Num_Prior_Violent = factor(RA2_Num_Prior_Violent, levels=c("1","2","3"), labels=c(0,1,2)),
    RA2_How_Many_FTA_Last_2_Yrs = factor(RA2_How_Many_FTA_Last_2_Yrs, levels=c("1","2","3"), labels=c(0,1,2)), # Questions don't match documentation exactly
    RA2_Has_Prior_Incaceration = factor(RA2_Has_Prior_Felony, levels=c("1","2"), labels=c(2,0))
  ) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric) 

NCA$score = apply(NCA,1,sum)
```


```{r}
# New Violent Criminal Activity (NCA)
NVCA = data_int %>%
  transmute(
    RA2_Current_Charge_Violent = factor(RA2_Current_Charge_Violent, levels=c("1","2"), labels=c(2,0)),
    RA2_Violent_and_under_21 = factor(RA2_Violent_and_under_21, levels=c("1","2"), labels=c(1,0)),
    RA2_New_Charge_While_Case_Pending = factor(RA2_New_Charge_While_Case_Pending, levels=c("1","2"), labels=c(1,0)),
    RA2_Has_Prior_Felony_Misdemeanor = factor(RA2_Has_Prior_Felony_Misdemeanor, levels=c("1","2"), labels=c(1,0)),
    RA2_Num_Prior_Violent = factor(RA2_Num_Prior_Violent, levels=c("1","2","3"), labels=c(0,1,2))
  ) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

NVCA$score = apply(NVCA,1,sum)
```


```{r}
bind_cols(
  FTA=select(FTA,score),
  NCA=select(NCA,score),
  NVCA=select(NVCA,score)
)
```

```{r}
df = as.tbl(data.frame(
  FTA$score,
  NCA$score,
  NVCA$score
))

df$all_na = apply(df,1,function(x) all(is.na(x)))
df$any_na = apply(df,1,function(x) any(is.na(x)))

df2 = df %>%
  filter(!any_na) %>%
  select(FTA.score,NCA.score,NVCA.score) %>%
  mutate_all(as.factor)

ggpairs(df2)
```





## 03 CNet_Cases 

```{r}
data_cnet_cases = read_delim(paste0(path_data,'03_KY_Recidivism_CNet_Cases.txt'), '|', col_names=TRUE) %>%
  mutate_if(is.character, str_trim)

data_cnet_cases
```

```{r}
# Each row is a unique SeqCaseID
length(unique(data_cnet_cases$SeqCaseID)) == nrow(data_cnet_cases)
```



## 04 PRIM_Cases 

```{r}
data_prim_cases = read_delim(paste0(path_data,'04_KY_Recidivism_PRIM_Cases.txt'), '|', col_names=TRUE) %>%
  mutate_if(is.character, str_trim)

data_prim_cases
```

```{r}
# Each row is a unique CaseID (but SeqCaseID are repeated)
length(unique(data_prim_cases$CaseID)) == nrow(data_prim_cases)
```

## 06 CNet_Charges

```{r}
data_cnet_charges = read_delim(paste0(path_data,'06_KY_Recidivism_CNet_Charges.txt'), '|', col_names=TRUE) %>%
#data_cnet_charges = read_delim(paste0(path_data,'06_KY_Recidivism_CNet_Charges.txt'), '|', col_names=TRUE, n_max = 100) %>%
  mutate_if(is.character, str_trim)

data_cnet_charges
```


## 07 PRIM_Charges

```{r}
data_prim_charges = read_delim(paste0(path_data,'07_KY_Recidivism_PRIM_Charges.txt'), '|', col_names=TRUE) %>%
  mutate_if(is.character, str_trim)

data_prim_charges
```

## Comparing tables

The dates for SeqCaseID==3993 are confusing to me. Looks like the offense happened in 1995 according to cnet but 2011 in prim (although there are no charges, just the case). I thought having the same SeqCaseID meant they were the same case?

```{r}
x = 3993

cnet_case_x = filter(data_cnet_cases,SeqCaseID==x)
prim_case_x = filter(data_prim_cases,SeqCaseId==x)

cnet_charge_x = filter(data_cnet_charges,SeqCaseID==x)
prim_charge_x = filter(data_prim_charges,SeqCaseID==x)

View(cnet_case_x)
View(prim_case_x)

View(cnet_charge_x)
View(prim_charge_x)
```

