---
title: "PRIM"
author: "Bin Han"
date: "May 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(data.table)
```


### Load PRIM data

**Specify Path Before Proceeding**


```{r}
## defendants
Defendants = as.data.frame(
  fread(paste0(data_path,'01_KY_Recidivism_Defendants.txt'), header = T)
  ) %>% 
  select(PersonID, Gender)

## PRIM interviews
Interviews = as.data.frame(
  fread(paste0(data_path,'02_KY_Recidivism_Interviews.txt'))
  ) %>% 
  rename(PersonID = PersonId, 
         InterviewID = InterviewId, 
         Interview_Date_Created = DateCreated)

## PRIM cases
PRIM_cases = as.data.frame(
  fread(paste0(data_path,'04_KY_Recidivism_PRIM_Cases.txt'))
  ) %>%
  rename(SeqCaseID = SeqCaseId, 
         Case_Date_Booked = date_booked, 
         Age_at_Case_Booked = Age_at_Booking) %>%
  filter(County != "WHITLEY")


## PRIM charges
PRIM_charges = as.data.frame(
  fread(paste0(data_path,'07_KY_Recidivism_PRIM_Charges.txt'))
  ) %>%
  rename(InterviewID = interviewID, 
         PersonID = PersonId, 
         CaseID = CaseId, 
         Charge_DispositionDate = DispositionDate) %>% 
  filter(County != "WHITLEY")


## PRIM bonds
Bonds = as.data.frame(
  fread(paste0(data_path,'09_KY_Recidivism_Bonds.txt'))
  ) %>% 
  rename(InterviewID = interviewID, 
         PersonID = PersonId, 
         CaseID = CaseId)%>% 
  filter(County != "WHITLEY")

## PRIM events
Events = as.data.frame(
  fread(paste0(data_path,'10_KY_Recidivism_Events.txt'))
  ) %>% 
  rename(InterviewID = InterviewId,
         PersonID = PersonId, 
         CaseID = Caseid,
         SeqCaseID = SeqCaseid) %>% 
  filter(EventCounty != "WHITLEY")

## PRIM fta
FTA = as.data.frame(
  fread(paste0(data_path,'11_KY_Recidivism_FTA.txt'))
  ) %>% 
  rename(InterviewID = InterviewId, 
         PersonID = PersonId, 
         CaseID = Caseid, 
         SeqCaseID = SeqCaseid)

## supervision
Supervision = as.data.frame(
  fread(paste0(data_path,'12_KY_Recidivism_Supervision.txt'))
  ) %>% 
  rename(InterviewID = InterviewId, 
         PersonID = PersonId, 
         CaseID = Caseid)
```





### 1. Defendants 

```{r}
colnames(Defendants)
```

```{r}
summary(Defendants)
```

```{r}
length(unique(Defendants$PersonID)) == nrow(Defendants)
```

__NOTE__:

Eahc row has an unique PersonID, representing a unique defendant.





## 2. Interview 

```{r}
colnames(Interviews)
```

```{r}
Interviews = Interviews %>% 
    select(-c(HoldingCountyCode, 
              County, 
              RiskLevel, 
              RiskScore, 
              HasHolder, 
              PostedPrior))
```

```{r}
summary(Interviews)
```

```{r}
length(unique(Interviews$PersonID)) == nrow(Interviews)
Interviews %>% group_by(PersonID) %>% summarise(count = n()) %>% arrange(desc(count))
Interviews %>% group_by(PersonID, InterviewID) %>% summarise(count = n()) %>% arrange(desc(count))
```

__NOTE__:

Each person (PersonID) could correspond to several interviews. But the combination of PersonID and InterviewID is unique.

```{r}
Interviews[Interviews$PersonID == "28566", ]
```







## 3. Cases 

```{r}
colnames(PRIM_cases)
```

```{r}
summary(PRIM_cases)
PRIM_cases[is.na(PRIM_cases$InterviewID), ]
```

```{r}
hist(PRIM_cases$Age_at_Case_Booked)
sort(unique(PRIM_cases$Age_at_Case_Booked))
sum(PRIM_cases$Age_at_Case_Booked >= 80)
```

__NOTE__:
1. There are 3 NA's on InterviewID, CaseID, Age_at_Case_Booked. Same three. Error.
2. Max age_at_case_booked: 237. And there are 6369 people with age at case booked > 80. Typos.

```{r}
PRIM_cases = PRIM_cases %>% 
    select(-c(CountyCode, 
              AlternativeReleaseType, 
              AlternativeReleaseDate, 
              Mcr, 
              McrDate)) %>% 
    filter(!is.na(InterviewID))
```


```{r}
length(unique(PRIM_cases$CaseID)) == nrow(PRIM_cases)
length(unique(PRIM_cases$SeqCaseID)) == nrow(PRIM_cases)
```

__NOTE__:
Each row is an unique case with CaseID. But SeqCaseID has some repeated appearance. It indicates that each SeqCaseID could relate to several CaseID's.

## sanity check
## same person, same interview
```{r}
PRIM_cases %>% group_by(PersonID, InterviewID) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
PRIM_cases[PRIM_cases$PersonID == "110149" & PRIM_cases$InterviewID == "686821", ]
```

## same person, same interview, same case
```{r}
PRIM_cases %>% group_by(PersonID, InterviewID, CaseID) %>% summarise(count = n()) %>% arrange(desc(count))
```

## same person, same interview, same seqcase
```{r}
PRIM_cases %>% group_by(PersonID, InterviewID, SeqCaseID) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
PRIM_cases[PRIM_cases$PersonID == "409241" & PRIM_cases$InterviewID == "1375331" & PRIM_cases$SeqCaseID == "1719208", ]
```

## same person, same seqcase
```{r}
PRIM_cases %>% group_by(PersonID, SeqCaseId) %>% summarise(count = n()) %>% arrange(desc(count))
```

```{r}
PRIM_cases[PRIM_cases$PersonID == "152948" & PRIM_cases$SeqCaseId == "913837", ]
```


__Insights__:
1. Same person, same interview, could correspond to the different SeqCaseID.
2. Same person, same interview, same SeqCaseID, could have different CaseIDs.
3. Same person, same SeqCaseID, could correspond to multiple interview.





## 4. Charges

```{r}
colnames(PRIM_charges)
```

```{r}
summary(PRIM_charges)
```




## 5.Bonds
```{r}
colnames(Bonds)
```

```{r}
summary(Bonds)
```







### Combine Data Sets

## 1. defendants + interviews
```{r}
comb1 = merge(x=Defendants, y=Interviews, by="PersonID")
nrow(comb1) == nrow(Interviews)
```

__NOTE__:
No information lost or added. 

```{r}
colnames(comb1)
```






## 2. comb1 + PRIM_cases
```{r}
## inner join
comb2 = merge(x=comb1, y=PRIM_cases, by=c("PersonID", "InterviewID"))
nrow(comb2) == nrow(PRIM_cases)
```

__NOTE__:
No information added or lost.


```{r}
colnames(comb2)
```





## comb2 + PRIM_charges
```{r}
comb3 = merge(x=comb2, 
              y=PRIM_charges,
              by = c("PersonID", "InterviewID", "CaseID", "SeqCaseID"))
```

```{r}
a = comb3 %>% 
    group_by(PersonID, SeqCaseID, InterviewID, CaseID) %>% 
    summarise(count = n()) %>% 
    arrange(desc(count))
b = a[a$count>1, ]
sum(b$count - 1) == (nrow(comb3) - nrow(comb2))
```

__NOTE__: Extra records added. It is because each case could have multiple charges. Sanity check conducted above.

```{r}
colnames(comb3)
```

```{r}
comb3 = comb3 %>% 
    select(-County.y,
           -Division.y,
           -CaseNumberType.y) %>% 
    mutate(County = County.x,
           Divison = Division.x,
           CaseNumberType = CaseNumberType.x)
```



```{r}
PRIM_cases = comb2
save(PRIM_cases, file = "PRIM_Cases.RData")
```


