---
title: "KY Features"
author: "Bin Han"
date: "June 23, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(purrr)
library(stringr)
library(magrittr)
library(lubridate)
library(readr)
library(ggplot2)
```

```{r}
load("CNet.RData")
load("PRIM_Cases.RData")
load("PRIM_Bonds.RData")
```


## CourNet Data

## 1. Pre-processing

Get ride of some redundant variables that are not necessary for prediction (add back when needed). Since we will merge this CourtNet data set with the PRIM data set based on case level, we get rid of CourNet cases that have no appearance in the PRIM system. If the value of "PRIM" is 0, the case is excluded from the data. We only included UOR levels that are "misdemeanor", "felony", and "other".

- Case_FilingDate: CourtNet case filing date
- Age_at_Casing Filing: age at courtnet case file
- ChargeDate: courtnet charge date
- UOR_Description: charge description
- UOR_Level: charge level
- sentence: category of sentence
- SentMonths: total prison or jail time of a sentence in months


Notice that CNet data could date back before 1/1/1996, and is reliable until this time. So to make sure that our data has better quality, case filing date before 1/1/1996 are excluded. 

Additionally, since people with "SentMonth" larger than 24 months (2 years) definitely cannot commit crime again within two years (at least publicly, maybe in jail, but out of our scope of consideration). Therefore, for those people who have been sentenced before, we included individuals only with "SentMonth" less than or equal to 12 months (filtering process in later step). We want to make sure that individuals that have been sentenced at the screening date (or current offense date) have at least one year out of two years to be in public. Then we see if they commited crime again within two years after they are released.

Some specific notes of the following step:

 - UOR_Level: only include charges with misdemeanor, felony, and other level. Excluded "violation" level of charges, which mostly consist of some traffic violations.
 - CNet_case_filing_date > "1996-01-01": get reliable records by excluding outdated ones
 - Filtered charge dispositions with code of "ACQ", "CAD", "DEC", "DGJ", "DIS","MST", "NG", "NGJ","PRDIS", "WD" (convicted charged).
 - Sentence Month >= 0: some of the recorded sentence month values are negative, whhich are invalid.
 -Sentence Month == NA: some people do not have sentence information, which are not entered in the system and they show up as NA


```{r}
CNet = CNet_data %>% select(PersonID, 
                            SeqCaseID, 
                            Gender, 
                            Race, 
                            PRIM, 
                            Case_FilingDate, 
                            Age_at_Case_Filing,
                            ChargeEntityNumber, 
                            ChargeDate, 
                            Charge_DispositionCode,
                            UOR_Description, 
                            UOR_Level, 
                            UOR_Code,
                            sentence, 
                            SentMonths, 
                            SuspSentMonths, 
                            CondDisSentMonths,
                            ADE, 
                            Undergo_Treat) %>%  
    mutate(CNet_Case_Filing_Date = as.Date(Case_FilingDate),
           Age_at_CNet_Case_Filing = Age_at_Case_Filing,
           UOR_Description = toupper(UOR_Description),
           UOR_Code = as.numeric(as.character(UOR_Code)),
           ChargeDate = as.Date(ChargeDate),
           SentMonths = as.numeric(as.character(SentMonths)),
           CondDisSentMonths = as.numeric(as.character(CondDisSentMonths)),
           SuspSentMonths = as.numeric(as.character(SuspSentMonths)),
           ADE = ifelse(ADE == "1", 1, ifelse(ADE == "NA", NA, 0)),
           Undergo_Treat = ifelse(Undergo_Treat == "1", 1, 
                           ifelse(Undergo_Treat == "NA", NA, 0))) %>% 
    filter(UOR_Level %in%  c("MISDEMEANOR", "FELONY", "OTHER"),
           CNet_Case_Filing_Date >= "1996-01-01",
           !is.na(ChargeDate),
           !(Charge_DispositionCode %in% 
               c("ACQ", "CAD", "DEC", "DGJ", "DIS","MST", "NG", "NGJ","PRDIS", "WD")),
           SentMonths >= 0 | is.na(SentMonths)) %>% 
    select(-c(PRIM, 
              Case_FilingDate, 
              Age_at_Case_Filing, 
              Charge_DispositionCode))
```



```{r}
summary(CNet)
```


__NOTE__:
The age variable is not right. There are so many typos. Check PRIM cases later since there is also one age variable in the PRIM_cases data set. Check if the two data sets have the same issue and determine which one is relatively more accurate. 

The sentence month is not right either. 114 records have sentence month greater than 600 months, which is equal to 50 years. Based on some external information that "the Class A felony has max 50 years sentence", those extremely large sentences should be typos. They will be filtered out in later steps. 


## 2. Add basic features

- is_violent
- is_felony
- is_misdemeanor
- is_property
- is_murder
- is_assault
- is_sex_offense
- is_weapon
- is_fel_prop_violence
- is_fel_assult
- is_misdemeanor_assault
- is_traffic
- is_drug
- is_dui
- is_domestic_violence
- is_stalking
- is_voyeurism
- is_fraud
- is_stealing/theft
- is_trespass

__NOTE__:

- The Case_NumberType does not match with the UOR charge level. Decided to proceed with the UOR charge level.
- It is not very precise to simply use key word to find what type of crime the charge is involved in. So we checked the code range for each type of crime (may still miss some charges)


## use uor_code + uor_description
```{r}
CNet_features = CNet %>% 
    mutate(
        is_violence = ifelse(str_detect(UOR_Description, "VIOLENCE") | 
                             UOR_Code %in% seq(130130, 130325), 1, 0),
        
        is_felony = ifelse(UOR_Level == "FELONY", 1, 0),
        
        is_misdemeanor = ifelse(UOR_Level == "MISDEMEANOR", 1,0),
        
        is_property = ifelse(str_detect(UOR_Description, "PROPERTY"), 1, 0),
        
        is_murder = ifelse(str_detect(UOR_Description, "MURDER") | 
                           UOR_Code %in% c(seq(90701, 90705), seq(91000, 91005), 
                                           seq(91500, 91605), seq(92010, 92025),
                                           seq(92100, 93035), seq(99900, 99905)), 1, 0),
        
        is_assault = ifelse(str_detect(UOR_Description, "ASSAULT"), 1, 0),
        
        is_sex_offense = ifelse(UOR_Code %in% c(seq(16000, 16605), seq(16670, 17245), 
                                seq(100300, 100405), seq(109900, 112305)), 1, 0),
        
        is_weapon = ifelse(str_detect(UOR_Description, "WEAPON") | 
                           str_detect(UOR_Description, "GUN") |
                           str_detect(UOR_Description, "FIREARM") |    
                           UOR_Code %in% c(seq(15010, 15065), seq(90701, 90705)), 1, 0),
        
        is_fel_prop_viol = ifelse(is_felony == 1 & is_property == 1 & is_violence == 1, 1, 0),
        
        is_fel_assult = ifelse(is_felony == 1 & is_assault==1, 1, 0),
        
        is_misde_assult = ifelse(is_misdemeanor == 1 & is_assault==1, 1, 0),
        
        is_traffic = ifelse(str_detect(UOR_Description, "TRAFFIC") | 
                            UOR_Code %in% c(seq(10,1460), seq(1500, 4325), 
                                            seq(4380, 4405), seq(4420, 7295)) , 1,0),
        
        is_drug = ifelse(str_detect(UOR_Description, "DRUG") | 
                         UOR_Code %in% c(seq(18020,18915), seq(48000, 48005), 
                                         seq(418060, 418216), seq(419900, 419995),
                                         seq(420090,420095),
                                         seq(421010, 421166), seq(421310, 421626)), 1, 0),
        
        is_dui = ifelse(str_detect(UOR_Description, "DUI") | 
                        UOR_Code %in% c(seq(1470,1495), seq(26190, 26205), 
                                        seq(26281, 26295)), 1, 0),
        
        is_stalking = ifelse(str_detect(UOR_Description, "STALKING") | 
                             UOR_Code %in% seq(27640, 27645), 1, 0),
        
        is_voyeurism = ifelse(str_detect(UOR_Description, "VOYEURISM") | 
                              UOR_Code %in% seq(17300, 17405), 1, 0),
        
        is_fraud = ifelse(str_detect(UOR_Description, "FRAUD") | 
                          UOR_Code %in% seq(16650, 16655), 1, 0),
        
        is_stealing = ifelse(str_detect(UOR_Description, "THEFT") |
                             str_detect(UOR_Description, "ROBBERY") | 
                             UOR_Code %in% seq(120000, 122025), 1, 0),
        
        is_trespass = ifelse(str_detect(UOR_Description, "TRESPASS") | 
                             UOR_Code %in% seq(26160, 26175), 1, 0),
        
        jail = ifelse(is.na(sentence), 0, ifelse(sentence == "Jail Time", 1, 0)),
        
        prison = ifelse(is.na(sentence), 0, ifelse(sentence == "Prison Time", 1, 0)),
        
        SentMonths = ifelse(is.na(SentMonths), 0, SentMonths),
        
        SuspSentMonths = ifelse(is.na(SuspSentMonths), 0, SuspSentMonths),
        
        CondDisSentMonths = ifelse(is.na(CondDisSentMonths), 0, CondDisSentMonths),
        
        ADE = ifelse(is.na(ADE), 0, ifelse(ADE == 1, 1, 0)),
        
        Treatment = ifelse(is.na(Undergo_Treat),0, ifelse(Undergo_Treat == 1, 1, 0)),
        
        incarceration = ifelse(jail == 1|prison == 1, 1, 0)
    ) %>% 
  select(-c(UOR_Description, 
            UOR_Level, 
            UOR_Code, 
            sentence, 
            Undergo_Treat, 
            jail, 
            prison)) 
```


```{r}
summary(CNet_features)
```




###### PRIM Cases ##########################################################################

## 1. preprocessing

Get ride of some redundant variables that are not necessary for prediction (add back when needed). The "Fta" can be used to determine if the person failed to appear for this case or not. "RA1" features are convenient variables as criminal history.

The interview date starts from 01/01/2009 in the PRIM system. However, the risk scores are only effective after 07/01/2013. Therefore, to test PSA, we need to restrict the interview date to be on and after 07/01/2013.

```{r}
PRIM_features0 = PRIM_cases %>% 
  select(PersonID, 
         SeqCaseID, 
         CaseID, 
         Interview_Date_Created,
         Case_Date_Booked, 
         Age_at_Case_Booked,
         OnProbation, 
         SupProbation, 
         Fta, 
         fta_risk_score_raw, 
         nca_risk_score_raw, 
         pvf_risk_score_raw, 
       
         ## risk score factors
         RA2_Current_Charge_Violent,
         RA2_Violent_and_under_21, 
         RA2_Age_At_Interview,
         RA2_New_Charge_While_Case_Pending,
         RA2_Has_Prior_Misdemeanor, 
         RA2_Has_Prior_Felony,
         RA2_How_Many_FTA_Last_2_Yrs,
         RA2_Has_Any_FTA_Older_2_Yrs,
         RA2_Num_Prior_Violent,
         RA2_Has_Prior_Incarceration) %>% 
  mutate(FTA = ifelse(Fta == "Y", 1, 0), ## case level
         probation = ifelse(OnProbation == "Y"|SupProbation == "Y", 1, 0), ## case level
         Interview_Date = as.Date(Interview_Date_Created),
         PRIM_Case_Filing_Date = as.Date(Case_Date_Booked),
         Age_at_PRIM_Case_Filing = Age_at_Case_Booked,
         fta_risk_score_raw = ifelse(fta_risk_score_raw == "NULL", NA, fta_risk_score_raw),
         nca_risk_score_raw = ifelse(nca_risk_score_raw == "NULL", NA, nca_risk_score_raw),
         pvf_risk_score_raw = ifelse(pvf_risk_score_raw == "NULL", NA, pvf_risk_score_raw)) %>%
  filter(Interview_Date >= "2013-07-01") %>%  ## to ensure that Arnold Scores are valid
  select(-c(Fta,OnProbation, 
            SupProbation, 
            Case_Date_Booked, 
            Age_at_Case_Booked, 
            Interview_Date_Created))
```

```{r}
summary(PRIM_features0)
```


```{r}
PRIM_features0 %>% group_by(PersonID, SeqCaseID) %>% summarise(count = n ()) %>% arrange(desc(count))
```

__NOTE__:
One case in CourtNet should be matched exactlt to one case in PRIM, using SeqCaseID. The case booked date in PRIM system should be slightly later than the case booked date in the CourtNet system. However, there are multiple reasons why there are many caseIDs corresponding to one SeqCaseID in the PRIM system:

1. for the original case, the defandent may fail to appear. Then the defendant will be arrested again and a new interview and a new case will be created. The new CaseID in the PRIM system is still under the same SeqCaseID from the original case. 
   - PersonID == 5; SeqCaseID == 3046161

2. There are cases when multiple CaseID's are related to one SeqCaseID, which means that there should be more than one pretrial interview related to that one SeqCaseID based on the explanation. However, the interviewID is the same for all CaseID's. 

3. The reason why we saw different risk scores for the child CaseID's of the same SeqCaseID is that if the FTA appears in the previous case, which is one the factors to calculate risk score, the risk score will definitely change. 

4. PRIM_Cases data set has the same age issue, that is the age at PRIM case filing ranges from 0 to 237, which is impossible. Need to compare the two data sets to see which one is relatively more accurate


```{r}
PRIM_features = PRIM_features0 %>% 
  select(PersonID, 
         SeqCaseID, 
         CaseID, 
         Interview_Date,
         FTA, 
         probation, 
         RA2_New_Charge_While_Case_Pending,
         RA2_Current_Charge_Violent, ## charge level
         RA2_Has_Prior_Incarceration,
         PRIM_Case_Filing_Date, 
         Age_at_PRIM_Case_Filing,
         fta_risk_score_raw, 
         nca_risk_score_raw, 
         pvf_risk_score_raw) %>% 
  mutate(pending_charge = ifelse(RA2_New_Charge_While_Case_Pending == "2", 0, 
                          ifelse(RA2_New_Charge_While_Case_Pending == "1", 1, NA)),
         is_violence_prim = ifelse(RA2_Current_Charge_Violent == "2", 0, 
                            ifelse(RA2_Current_Charge_Violent == "1", 1, NA)),
         
         prior_incar_prim = ifelse(RA2_Has_Prior_Incarceration == "2", 0,
                                   ifelse(RA2_Has_Prior_Incarceration == "1", 1, NA)),
         
         fta_risk_score_raw = as.numeric(fta_risk_score_raw),
         nca_risk_score_raw = as.numeric(nca_risk_score_raw),
         pvf_risk_score_raw = as.numeric(pvf_risk_score_raw)) %>% 
  select(-RA2_New_Charge_While_Case_Pending, 
         -RA2_Current_Charge_Violent, 
         -RA2_Has_Prior_Incarceration) %>%
  na.omit(.)
```


```{r}
summary(PRIM_features)
```




## Combine PRIM_CASES + CNet_features

Merge CourNet features and PRIM features. Get rid of null charge date and change the formate to date format.

```{r}
all_features0 = merge(x=CNet_features, y=PRIM_features, by=c("PersonID", "SeqCaseID"))
```


## check violence feature
```{r}
violence_check = all_features0 %>% 
  select(PersonID, 
         SeqCaseID, 
         ChargeEntityNumber, 
         ChargeDate,is_violence, 
         is_violence_prim)

mean(violence_check$is_violence)
mean(violence_check$is_violence_prim)
```

__NOTE__:

By checking the violence feature from CourtNet and from PRIM, PRIM has more records of 1 when CNet has value of 0. It makes sense since when checking if a charge is violent or not in CourtNet system, we detect a certain string in UOR Description. But PRIM directly has the answer from interview. Therefore, we decided to proceed with PRIM feature.

```{r}
all_features0 = all_features0 %>% 
  select(-is_violence) %>% 
  rename(is_violence = is_violence_prim)
```


## sanity check
```{r}
a = all_features0[all_features0$PersonID == "5" & all_features0$SeqCaseID == "3046161", ]
b = PRIM_features[PRIM_features$PersonID == "5"& PRIM_features$SeqCaseID == "3046161",]
c = CNet[CNet$PersonID == "5" & CNet$SeqCaseID == "3046161", ]
```


__NOTE__:
There are extra records added to the data set. It is because one SeqCaseID could relate to multiple CaseID's in the PRIM system. Additionally, one SeqCaseID could have multiple charges. Therefore, each one of the CaseIDs will have the same number of charges information added.

Take the example above for instance, the individual with PersonID == 5 & SeqCaseID == 3046161 had 5 charges under this CourtNet case. This CourtNet case (SeqCaseID) had 2 CaseIDs corresponded in PRIM system. Therefore, after inner joining them, there are 10 records in total, since each CaseID will have 5 charges right now. The 5 charges are exactly the same in both CaseIDs. `This is very important since it will change the way we calcualted features later on.`



##### age issue 
```{r}
age_issue = all_features0 %>% select(PersonID, 
                                     SeqCaseID, 
                                     CaseID, 
                                     CNet_Case_Filing_Date, 
                                     Age_at_CNet_Case_Filing, 
                                     PRIM_Case_Filing_Date,
                                     Age_at_PRIM_Case_Filing, 
                                     ChargeDate) 
sum(age_issue$Age_at_CNet_Case_Filing < 18 | age_issue$Age_at_CNet_Case_Filing > 70)
sum(age_issue$Age_at_PRIM_Case_Filing < 18 | age_issue$Age_at_PRIM_Case_Filing > 70)
```


Checked the two sides of both "age_at_CNet_case_filing" and "age_at_PRIM_case_filing". There are more values of CNet age lying within the normal range.  Therefore, to get each individual's DOB and age at each charge date we proceed with CNet_Case_Filing_Date and Age_at_CNet_Case_Filing Date. To be consistent with the analysis in Broward, we limited the age range to be witin 18 and 70. Additionally, since different PRIM_case_filing_date could have the same filing_age, there might be different DOBs for one person. (PersonID = 8, SeqCaseID = 2523727)


```{r}
all_features = all_features0 %>% 
    mutate(DOB = CNet_Case_Filing_Date - years(Age_at_CNet_Case_Filing),
           age_at_charge = floor(as.numeric(as.period(interval(DOB, ChargeDate)), "years"))) %>%
    filter (age_at_charge >= 18,
            age_at_charge <= 70) %>% 
    select(-c(CNet_Case_Filing_Date, 
              Age_at_CNet_Case_Filing,
              Age_at_PRIM_Case_Filing, DOB,
              Interview_Date))
```


```{r}
summary(all_features)
```










## Split the combined data 
## Find cutoff date and screening date

- Cutoff Date: three years eariler than the latest charge date in our data set

The reason we use 3 years here is that for those people who are sentenced to jail/prison for 1 year or less, we'll use the date when they are released and see if they recidivated or not. We want to make sure that everyone does have the same two-year period to be checked.


```{r}
## cutoff date
cutoff_date = max(all_features$ChargeDate) - years(3)

## subset data
before_cutoff0 = all_features %>% filter(ChargeDate < cutoff_date)
after_cutoff0 = all_features %>% filter(ChargeDate > cutoff_date)
```

- Screening Dates: for each individual, the cloesest charge date to the cutoff date will be used as the "screening date". Before and on the screening date, the information will be used as criminal history. After the screening date, the information will be used to determine recidivism. The screening date is also referred to as "current charge date"


```{r}
## screening dates
screening_dates = before_cutoff0 %>% 
    group_by(PersonID) %>% 
    summarise(screening_date = max(ChargeDate))

## merge screening dates with before_cutoff data set
before_cutoff0 = merge(x=before_cutoff0, y=screening_dates, by="PersonID", all.x = T) %>% 
    mutate(Gender = ifelse(Gender == "Male", 1, 0))
```

#### Add year indicator features

Since we have charge dates and screening dates, now we can calculate the time interval between each charge date and the screening date and then determine whether or not the person had committed crime in the past 5 years, 3 years, 1 year, and 6 months.

The year_case variable is for each PRIM case, since the FTA and probation information is based on PRIM case level. Using the year_case, we can determine the other two features: "fta_two_year" and "fta_two_year_plus".

```{r}
before_cutoff = before_cutoff0 %>% 
    mutate(year_charges = as.numeric(as.period(interval(ChargeDate, screening_date)), "years"),
           year_cases = as.numeric(as.period(interval(PRIM_Case_Filing_Date, 
                                                      screening_date)), "years"),
           
           ## fta information is based on case level
           fta_two_year = ifelse(FTA == 1 & year_cases <= 2, 1, 0),
           fta_two_year_plus = ifelse(FTA == 1 & year_cases > 2, 1, 0),
           
           ## had charges within specific time frame
           six_month = ifelse(year_charges <= 0.5, 1, 0),
           one_year = ifelse(year_charges <= 1, 1, 0),
           three_year = ifelse(year_charges <= 3, 1, 0),
           five_year = ifelse(year_charges <= 5, 1, 0)) %>% 
    select(-c(FTA, year_charges, year_cases, PRIM_Case_Filing_Date))
```

```{r}
summary(before_cutoff)
```







### function to create features
```{r}
create_features = function(data){

    ## first layer of group: CaseID level. 
  
    features = data %>% 
        group_by(PersonID, screening_date, SeqCaseID, CaseID, Gender, Race) %>% 
        summarise(age_at_current_charge = max(age_at_charge),
                  p_SentMonths = max(SentMonths),
                  
                  ## charge entity number is in order
                  p_charges = max(ChargeEntityNumber)-min(ChargeEntityNumber)+1,
                  
                  p_violence = sum(is_violence), 
                  p_felony = sum(is_felony),
                  p_misdemeanor = sum(is_misdemeanor),
                  p_property = sum(is_property),
                  p_murder = sum(is_murder),
                  p_assault = sum(is_assault),
                  p_sex_offense = sum(is_sex_offense),
                  p_weapon = sum(is_weapon),
                  p_felprop_viol = sum(is_fel_prop_viol),
                  p_felassult = sum(is_fel_assult),
                  p_misdeassult = sum(is_misde_assult),
                  p_traffic = sum(is_traffic),
                  p_drug = sum(is_drug),
                  p_dui = sum(is_dui),
                  p_stalking = sum(is_stalking),
                  p_voyeurism = sum(is_voyeurism),
                  p_fraud = sum(is_fraud),
                  p_stealing = sum(is_stealing),
                  p_trespass = sum(is_trespass),
                  
                  ## Even though ADE and Treatment are sentence level (charge level) information,
                  ## as long as the individual has one ADE/Treatment value that is equal to 1,
                  ## the individual has ADE/Treatment = 1 at the CaseID level. Therefore, we use
                  ## max function. The same logic applies to other binary indicator variables.
                  
                  ADE = max(ADE),
                  Treatment = max(Treatment),
                  p_incarceration = max(prior_incar_prim),
                  
                  ## Those four are binary indicator variables. Same logic above applies here.
                  ## fta_two_year, fta_two_year_plus, probation are CaseID level information. 
                  ## We use max(). pending_charge is the charge level information. But as long as
                  ## the individual has pending charge at any one of the charges under one case,
                  ## the case then has pending_charge indicator as 1.
                  
                  p_fta_two_year = max(fta_two_year),
                  p_fta_two_year_plus = max(fta_two_year_plus),
                  p_probation = max(probation),
                  p_pending_charge = max(pending_charge),
                  
                  six_month = max(six_month),     ## as long as there is one charge that is
                  one_year = max(one_year),       ## within six month/one year/three years/five
                  three_year = max(three_year),   ## years, the indicator variable is one.
                  five_year = max(five_year),
                  
                  aronld_fta_raw = mean(fta_risk_score_raw),
                  aronld_nca_raw = mean(nca_risk_score_raw),
                  aronld_nvca_raw = mean(pvf_risk_score_raw)) %>% 
      
## second layer group: for criminal history information; SeqCaseID level.
        group_by(PersonID, screening_date, SeqCaseID, Gender, Race) %>% 
        summarise(p_arrest = length(CaseID),
                  
                  ## Here we use max() function because the CaseIDs under the same SeqCaseID 
                  ## share the same information. Therefore, we just need one of those values.
                  ## We do not use mean() since the it will give us some decimals values.
                  
                  p_SentMonths = max(p_SentMonths),
                  age_at_current_charge = max(age_at_current_charge),
                  p_charges = max(p_charges),
                  p_violence = max(p_violence),
                  p_felony = max(p_felony),
                  p_misdemeanor = max(p_misdemeanor),
                  p_property = max(p_property),
                  p_murder = max(p_murder),
                  p_assault = max(p_assault),
                  p_sex_offense = max(p_sex_offense),
                  p_weapon = max(p_weapon),
                  p_felprop_viol = max(p_felprop_viol),
                  p_felassult = max(p_felassult),
                  p_misdeassult = max(p_misdeassult),
                  p_traffic = max(p_traffic),
                  p_drug = max(p_drug),
                  p_dui = max(p_dui),
                  p_stalking = max(p_stalking),
                  p_voyeurism = max(p_voyeurism),
                  p_fraud = max(p_fraud),
                  p_stealing = max(p_stealing),
                  p_trespass = max(p_trespass),
                  ADE = max(ADE),
                  Treatment = max(Treatment),
                  p_incarceration = max(p_incarceration),
                  
                  ## CourtNet Case Level Information
                  p_fta_two_year = sum(p_fta_two_year),
                  p_fta_two_year_plus = sum(p_fta_two_year_plus),
                  p_probation = sum(p_probation),
                  p_pending_charge = sum(p_pending_charge),
                  six_month = max(six_month),
                  one_year = max(one_year),
                  three_year = max(three_year),
                  five_year = max(five_year),
                  aronld_fta_raw = max(aronld_fta_raw),
                  aronld_nca_raw = max(aronld_nca_raw),
                  aronld_nvca_raw = max(aronld_nvca_raw)) %>% 
        ## third layer: individual level; sum up all the SeqCaseID information; individual level.
        group_by(PersonID, Gender, Race, screening_date) %>%   
        summarise(age_at_current_charge = max(age_at_current_charge),
                  p_SentMonths = sum(p_SentMonths),
                  p_arrest = sum(p_arrest),
                  p_charges = sum(p_charges),
                  p_violence = sum(p_violence),
                  p_felony = sum(p_felony),
                  p_misdemeanor = sum(p_misdemeanor),
                  p_property = sum(p_property),
                  p_murder = sum(p_murder),
                  p_assault = sum(p_assault),
                  p_sex_offense = sum(p_sex_offense),
                  p_weapon = sum(p_weapon),
                  p_felprop_viol = sum(p_felprop_viol),
                  p_felassult = sum(p_felassult),
                  p_misdeassult = sum(p_misdeassult),
                  p_traffic = sum(p_traffic),
                  p_drug = sum(p_drug),
                  p_dui = sum(p_dui),
                  p_stalking = sum(p_stalking),
                  p_voyeurism = sum(p_voyeurism),
                  p_fraud = sum(p_fraud),
                  p_stealing = sum(p_stealing),
                  p_trespass = sum(p_trespass),
                  ADE = max(ADE),  
                  Treatment = max(Treatment),
                  p_incarceration = max(p_incarceration),
                  p_fta_two_year = sum(p_fta_two_year),
                  p_fta_two_year_plus = sum(p_fta_two_year_plus),
                  p_pending_charge = sum(p_pending_charge),
                  p_probation = sum(p_probation),
                  six_month = max(six_month),
                  one_year = max(one_year),
                  three_year = max(three_year),
                  five_year = max(five_year),
                  aronld_fta_raw = max(aronld_fta_raw),
                  aronld_nca_raw = max(aronld_nca_raw),
                  aronld_nvca_raw = max(aronld_nvca_raw))
    return(features)
}
```


## Before Screening Date 

- six_month, one-year, three_year, five_year: whether or not the person has committed crimes within the last six_month, one_year, three_year, and five_year respectively.
- p_SentMonths: total amount of time the person has been sentence to jail/prison.

```{r}
data_before = before_cutoff %>% filter(ChargeDate < screening_date)
features_before = create_features(data_before) %>% 
    select(PersonID, 
           Race, 
           Gender, 
           screening_date, 
           six_month, 
           one_year, 
           three_year, 
           five_year, 
           p_SentMonths)
```


```{r}
summary(features_before)
```


## On Screening Date

```{r}
data_on = before_cutoff %>% filter(ChargeDate == screening_date)
features_on = data_on %>% 
    group_by(PersonID) %>% 
    summarise(current_violence = max(is_violence), 
              current_pending_charge = max(pending_charge))
```

```{r}
summary(features_on)
```


## Before and On Screening Date 

The last step was to further improve the data quality. The p_SentMonths is the total amount of jail/prison time that each individual had before the current charge. Since our data set excluded juvnile cases, therefore, we would expect that the age at current offense to be equal to or larger than the privious sentenced years + 18 years. If not, then there must be something wrong with the sentence month information, and the record needs to be removed.

```{r}
## data
data_before_on = before_cutoff %>% filter(ChargeDate <= screening_date)

## get features
features_before_on = create_features(data_before_on) %>% 
  select(-c(six_month, 
            one_year, 
            three_year, 
            five_year, 
            p_SentMonths)) %>% 
  merge(x = ., 
        y = features_before, 
        by = "PersonID", 
        all.x = TRUE) %>% 
  ## if all the following variables are NA, it means that those individuals do not have any
  ## criminal history besides the current charges.
  mutate(six_month = ifelse(is.na(six_month), 0, six_month),
         one_year = ifelse(is.na(one_year), 0, one_year),
         three_year = ifelse(is.na(three_year), 0, three_year),
         five_year = ifelse(is.na(five_year), 0, five_year),
         p_SentMonths = ifelse(is.na(p_SentMonths), 0, p_SentMonths))%>% 
  select(-c(Gender.y, Race.y, screening_date.y)) %>% 
  rename(Gender = Gender.x,
         Race = Race.x,
         screening_date = screening_date.x) %>% 
  merge(x=., 
        y=features_on, 
        by = "PersonID", 
        all.x = TRUE) %>% 
  filter(age_at_current_charge >= p_SentMonths/12 + 18) %>% 
  mutate(current_violence20 = ifelse(current_violence == 1 & age_at_current_charge <= 20, 1, 0))
```


```{r}
summary(features_before_on)
```


## After Screening Date

Since the screening dates are the latest dates of charges for each individual before the cutoff date, therefore, data after screening date is the same as data after the cutoff date. From now on, we still start using "after_cutoff" data set to create different prediction labels.

```{r}
## inner join two data sets
## if there is no PersonID from screening dates that have match in the after_cutoff data set, then those PersonID's did not commit any crime after the cutoff_date, or the screening date.


## sentence month information on the screening date (current offense date). We need this information to determine the date that the person was released and then see if any crime has been commited two years after this date.

release_dates = data_on %>% 
    group_by(PersonID, 
             screening_date) %>% 
    summarise(current_SentMonth = max(SentMonths),
              current_SuspSentMonth = max(SuspSentMonths),
              current_CondDisMonth = max(CondDisSentMonths)) %>% 
    mutate(SentDays = ifelse((current_SentMonth>0 & current_SuspSentMonth==0)|
                               (current_SentMonth>0 & current_CondDisMonth ==0),
                             round(current_SentMonth*30.8333), 0),
           release_date = screening_date+SentDays) %>% 
    select(PersonID, 
           release_date, 
           current_SentMonth, 
           current_SuspSentMonth, 
           current_CondDisMonth)

## merge
after_cutoff = merge(x=after_cutoff0, 
                     y=release_dates, 
                     by="PersonID", 
                     all.y = TRUE) %>% 
    filter((current_SentMonth <= 12) | 
             (current_SentMonth > 12 & current_SuspSentMonth > 0)|
             (current_SentMonth > 12 & current_CondDisMonth > 0)) %>% 
    select(-c(current_SentMonth, 
              current_SuspSentMonth, 
              current_CondDisMonth))

## add year feature
data_after = after_cutoff[, c(1,41,2:40)] %>% 
    mutate(year_offenses = as.numeric(as.period(interval(release_date, ChargeDate)), "years"),
           general_two_year = ifelse(year_offenses <= 2, 1, 0),
           drug_two_year = ifelse(general_two_year == 1 & is_drug == 1, 1, 0),
           violent_two_year = ifelse(general_two_year == 1 & is_violence == 1, 1, 0),
           felony_two_year = ifelse(general_two_year == 1 & is_felony == 1, 1, 0),
           misdemeanor_two_year = ifelse(general_two_year == 1 & is_misdemeanor == 1, 1, 0),
           property_two_year = ifelse(general_two_year == 1 & is_property == 1, 1, 0),
           
           general_six_month = ifelse(year_offenses <= 0.5, 1, 0),
           drug_six_month = ifelse(general_six_month == 1 & is_drug == 1, 1, 0),
           violent_six_month = ifelse(general_six_month == 1 & is_violence == 1, 1, 0),
           felony_six_month = ifelse(general_six_month == 1 & is_felony == 1, 1, 0),
           misdemeanor_six_month = ifelse(general_six_month == 1 & is_misdemeanor == 1, 1, 0),
           property_six_month = ifelse(general_six_month == 1 & is_property == 1, 1, 0)) %>% 
    select(PersonID:ChargeDate, ChargeEntityNumber, year_offenses:property_six_month)
```



## Generate labels

Using max() function because as long as there is one charge that is within two years of the screening date, the person has recidivism. 

```{r}
outcomes = data_after %>% 
    group_by(PersonID) %>% 
    summarise(general_two_year = max(general_two_year), 
              drug_two_year = max(drug_two_year),
              violent_two_year = max(violent_two_year),
              felony_two_year = max(felony_two_year),
              misdemeanor_two_year = max(misdemeanor_two_year),
              property_two_year = max(property_two_year),
  
              general_six_month = max(general_six_month),
              drug_six_month = max(drug_six_month),
              violent_six_month = max(violent_six_month),
              felony_six_month = max(felony_six_month),
              misdemeanor_six_month = max(misdemeanor_six_month),
              property_six_month = max(property_six_month))
              
```

#### Combine criminal history data with labels 

```{r}
recid_labels = merge(x=features_before_on, y=outcomes, by = "PersonID") %>% 
    mutate(general_two_year = ifelse(is.na(general_two_year), 0,  general_two_year),
           drug_two_year = ifelse(is.na(drug_two_year), 0,  drug_two_year),
           violent_two_year = ifelse(is.na(violent_two_year), 0,  violent_two_year),
           felony_two_year = ifelse(is.na(felony_two_year), 0,  felony_two_year),
           misdemeanor_two_year = ifelse(is.na(misdemeanor_two_year), 0,  misdemeanor_two_year),
           property_two_year = ifelse(is.na(property_two_year), 0,  property_two_year), 
          
           general_six_month = ifelse(is.na(general_six_month), 0,  general_six_month),
           drug_six_month = ifelse(is.na(drug_six_month), 0,  drug_six_month),
           violent_six_month = ifelse(is.na(violent_six_month), 0,  violent_six_month),
           felony_six_month = ifelse(is.na(felony_six_month), 0,  felony_six_month),
           misdemeanor_six_month = ifelse(is.na(misdemeanor_six_month), 0,
                                          misdemeanor_six_month),
           property_six_month = ifelse(is.na(property_six_month), 0,  property_six_month))
```

```{r}
recid_labels = recid_labels %>% 
  select(-p_SentMonths) %>% 
  rename(sex = Gender, 
         person_id = PersonID,
         race = Race)
recid_labels = recid_labels[, c(1:33, 37:43, 34:36, 44:55)]
```

```{r}
summary(recid_labels)
```



#### Save data files
```{r}
save(data_before,features_before, 
     data_on, features_on,
     data_before_on, data_after, features_before_on, 
     outcomes, recid_labels,
     file = "kentucy_features.Rdata")
```


```{r, message=FALSE}
set.seed(816)
test_sample = sample(1:nrow(recid_labels), 0.1*nrow(recid_labels), replace = F)
train = recid_labels[-test_sample,]
test = recid_labels[test_sample,]
```

```{r}
write.csv(recid_labels, file = "C:/Users/binha/Documents/Duke/Cynthia Research/KY-analysis-mytrials/KY Recidivism/KY data/kentucky_data.csv", row.names = F)

write.csv(train, file = "C:/Users/binha/Documents/Duke/Cynthia Research/KY-analysis-mytrials/KY Recidivism/KY data/kentucky_train.csv", row.names = F)

write.csv(test, file = "C:/Users/binha/Documents/Duke/Cynthia Research/KY-analysis-mytrials/KY Recidivism/KY data/kentucky_test.csv", row.names = F)
```















