---
title: "Features"
author: "Bin Han"
date: "June 23, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(purrr)
library(stringr)
library(magrittr)
library(lubridate)
library(readr)
library(ggplot2)
```

```{r}
load("CNet.RData")
load("PRIM_Cases.RData")
load("PRIM_Bonds.RData")
```


###### CourNet Data ##############################################################################

## 1. Pre-processing

Get ride of some redundant variables that are not necessary for prediction (add back when needed). Since we will merge this CourtNet data set with the PRIM data set based on case level, we get rid of CourNet cases that have no appearance in the PRIM system. If the value of "PRIM" is 0, the case is excluded from the data. We only included UOR levels that are "misdemeanor", "felony", and "other".

- Case_FilingDate: CourtNet case filing date
- Age_at_Casing Filing: age at courtnet case file
- ChargeDate: courtnet charge date
- UOR_Description: charge description
- UOR_Level: charge level
- sentence: category of sentence
- SentMonths: total prison or jail time of a sentence in months


Notice that Daniel mentioned the CNet data could date back before 1/1/1996, and is reliable until this time. So to make sure that our data has better quality (in terms of reliability), case filing date before 1/1/1996 are excluded. 

Additionally, since people with "SentMonth" larger than 24 months (2 years) definitely cannot commit crime again within two years (at least publicly, maybe in jail, but out of our scope of consideration). Therefore, for those people who have been sentenced before, we included individuals only with "SentMonth" less than or equal to 12 months (filtering process in later step). We want to make sure that individuals that have been sentenced at the screening date (or current offense date) have at least one year out of two years to be in public. Then we see if they commited crime again within two years after they are released.

```{r}
CNet = CNet_data %>% select(PersonID, SeqCaseID, Gender, PRIM, 
                            Case_FilingDate, Age_at_Case_Filing,
                            ChargeEntityNumber, ChargeDate, Charge_DispositionCode,
                            UOR_Description, UOR_Level, UOR_Code,
                            sentence, SentMonths, SuspSentMonths, CondDisSentMonths,
                            ADE, Undergo_Treat) %>%  
    mutate(CNet_Case_Filing_Date = as.Date(Case_FilingDate),
           Age_at_CNet_Case_Filing = Age_at_Case_Filing,
           UOR_Description = toupper(UOR_Description),
           UOR_Code = as.numeric(UOR_Code),
           ChargeDate = as.Date(ChargeDate),
           SentMonths = as.numeric(SentMonths) - as.numeric(CondDisSentMonths),
           SuspSentMonths = as.numeric(SuspSentMonths),
           ADE = ifelse(ADE == "1", 1, ifelse(ADE == "NA", NA, 0)),
           Undergo_Treat = ifelse(Undergo_Treat == "1", 1, 
                           ifelse(Undergo_Treat == "NA", NA, 0))) %>% 
    filter(UOR_Level %in%  c("MISDEMEANOR", "FELONY", "OTHER"),
           Case_FilingDate >= "1996-01-01",
           !is.na(ChargeDate),
           !(Charge_DispositionCode %in% 
               c("ACQ", "CAD", "DEC", "DGJ", "DIS","MST", "NG", "NGJ","PRDIS", "WD")),
           SentMonths > 0 | is.na(SentMonths)) %>% 
    select(-c(PRIM, Case_FilingDate, Age_at_Case_Filing, CondDisSentMonths, 
              Charge_DispositionCode))
```


```{r}
summary(CNet)
```


__NOTE__:
The age variable is not right. There are so many typos. Check PRIM cases later since there is also one age variable in the PRIM_cases data set. Check if the two data sets have the same issue and determine which one is relatively more accurate. Since we need filter out those typos. If one data set has more accurate captures, it can reduce the number of records that we will filter out.

The sentence month is not right either. 114 records have sentence month greater than 600 months, which is equal to 50 years. Based on some articles I read, the Class A felony has max 50 years sentence. I did not filter the sentencemonth here. It will be filtered out in later steps. Additionall, when sentence is NA, SentenceMonth, ADE and Undergo_Treat is NA as well. 


## check age distribution at normal age range
```{r}
age = CNet %>% 
  filter(Age_at_CNet_Case_Filing <= 70 & Age_at_CNet_Case_Filing >= 18) %>% 
  select(Age_at_CNet_Case_Filing)

age_count = age %>% group_by(Age_at_CNet_Case_Filing) %>% summarise(count = n())
ggplot(data = age_count, aes(x=Age_at_CNet_Case_Filing, y=count)) + geom_bar(stat = "identity")
```


## 2. Add basic features

- is_violence
- is_felony
- is_misdemeanor
- is_property
- is_murder
- is_assault
- is_family_violence (None)
- is_sex_offense
- is_weapon
- is_fel_prop_violence
- is_fel_assult
- is_misdemeanor_assault
- is_traffic
- is_drug
- is_dui
- is_domestic_violence
- is_stalking
- is_voyeurism
- is_fraud
- is_stealing/theft
- is_trespass

__NOTE__:

- The Case_NumberType does not match with the UOR charge level. Decided to proceed with the UOR level.
- It is not very precise to simply use key word to find what type of crime the charge is involved in. May need some future adjustment.


## use uor_code + uor_description
```{r}
CNet_features = CNet %>% 
    mutate(
        ## type of crime the charge is involved in
        is_violence = ifelse(str_detect(UOR_Description, "VIOLENCE") | 
                             UOR_Code %in% seq(130130, 130325), 1, 0),
        is_felony = ifelse(UOR_Level == "FELONY", 1, 0),
        is_misdemeanor = ifelse(UOR_Level == "MISDEMEANOR", 1,0),
        is_property = ifelse(str_detect(UOR_Description, "PROPERTY"), 1, 0),
        is_murder = ifelse(str_detect(UOR_Description, "MURDER") | 
                           UOR_Code %in% c(seq(90701, 90705), seq(91000, 91005), 
                                           seq(91500, 91605), seq(92010, 92025),
                                           seq(92100, 93035), seq(99900, 99905)), 1, 0),
        is_assault = ifelse(str_detect(UOR_Description, "ASSAULT"), 1, 0),
        is_sex_offense = ifelse(UOR_Code %in% c(seq(16000, 16605), seq(16670, 17245), 
                                seq(100300, 100405), seq(109900, 112305)), 1, 0),
        is_weapon = ifelse(str_detect(UOR_Description, "WEAPON") | 
                           str_detect(UOR_Description, "GUN") |
                           str_detect(UOR_Description, "FIREARM") |    
                           UOR_Code %in% c(seq(15010, 15065), seq(90701, 90705)), 1, 0),
        is_fel_prop_viol = ifelse(is_felony == 1 & is_property == 1 & is_violence == 1, 1, 0),
        is_fel_assult = ifelse(is_felony == 1 & is_assault==1, 1, 0),
        is_misde_assult = ifelse(is_misdemeanor == 1 & is_assault==1, 1, 0),
        is_traffic = ifelse(str_detect(UOR_Description, "TRAFFIC") | 
                            UOR_Code %in% c(seq(10,1460), seq(1500, 4325), 
                                            seq(4380, 4405), seq(4420, 7295)) , 1,0),
        is_drug = ifelse(str_detect(UOR_Description, "DRUG") | 
                         UOR_Code %in% c(seq(18020,18915), seq(48000, 48005), 
                                         seq(418060, 418216), seq(419900, 419995),
                                         seq(420090,420095),
                                         seq(421010, 421166), seq(421310, 421626)), 1, 0),
        is_dui = ifelse(str_detect(UOR_Description, "DUI") | 
                        UOR_Code %in% c(seq(1470,1495), seq(26190, 26205), 
                                        seq(26281, 26295)), 1, 0),
        is_stalking = ifelse(str_detect(UOR_Description, "STALKING") | 
                             UOR_Code %in% seq(27640, 27645), 1, 0),
        is_voyeurism = ifelse(str_detect(UOR_Description, "VOYEURISM") | 
                              UOR_Code %in% seq(17300, 17405), 1, 0),
        is_fraud = ifelse(str_detect(UOR_Description, "FRAUD") | 
                          UOR_Code %in% seq(16650, 16655), 1, 0),
        is_stealing = ifelse(str_detect(UOR_Description, "THEFT") |
                             str_detect(UOR_Description, "ROBBERY") | 
                             UOR_Code %in% seq(120000, 122025), 1, 0),
        is_trespass = ifelse(str_detect(UOR_Description, "TRESPASS") | 
                             UOR_Code %in% seq(26160, 26175), 1, 0),
        
        ## prison / jail sentence information
        jail = ifelse(is.na(sentence), 0, ifelse(sentence == "Jail Time", 1, 0)),
        prison = ifelse(is.na(sentence), 0, ifelse(sentence == "Prison Time", 1, 0)),
        SentMonths = ifelse(jail == 1 | prison == 1, SentMonths, 0),
        SuspSentMonths = ifelse(is.na(SuspSentMonths), 0, SuspSentMonths),
        ADE = ifelse(is.na(ADE), 0, ifelse(ADE == 1, 1, 0)),
        Treatment = ifelse(is.na(Undergo_Treat),0, ifelse(Undergo_Treat == 1, 1, 0))
    ) %>% ## further get rid of redundant variables
    select(-c(UOR_Description, UOR_Level, UOR_Code, sentence, Undergo_Treat)) %>% 
    filter(!is.na(SentMonths)) ## PersonID = 73272, SeqCaseID = 2674931: even "prison" ==
                               ## 1,SentMonth is still 0. Therefore, need to be removed.
```

```{r}
summary(CNet_features)
```


## drug population
```{r}
cols = colnames(CNet_features)
for (i in 11:29){
  print(i)
  column = cols[i]
  values = CNet_features[,c(1,2,10,i)] %>% 
    .[.[,4] == 1, ] %>% 
    filter(Age_at_CNet_Case_Filing <= 70,
           Age_at_CNet_Case_Filing >= 18) %>% 
    group_by(PersonID, SeqCaseID) %>% 
    summarise(age = max(Age_at_CNet_Case_Filing)) %>% 
    group_by(age) %>% 
    summarise(count = n())
  plot = ggplot(data = values, aes(x=age,y=count)) + geom_bar(stat = "identity") +
    xlab("age") + 
    ylab("count") + 
    ggtitle(column)
  print(plot)
}
```





###### PRIM Cases ##########################################################################

## 1. preprocessing

Get ride of some redundant variables that are not necessary for prediction (add back when needed). The "Fta" can be used to determine if the person failed to appear for this case or not. "RA1" features are cheat variables as criminal history.

The interview date starts from 01/01/2009 in the PRIM system. However, the risk scores are only effective after 07/01/2013. Therefore, to calculate the PSA scores test PSA, we need to restrict the interview date to be on and after 07/01/2013.

```{r}
PRIM_features0 = PRIM_cases %>% select(PersonID, SeqCaseID, CaseID, Interview_Date_Created,
                                      Case_Date_Booked, Age_at_Case_Booked,
                                      OnProbation, SupProbation, Fta, 
                                      fta_risk_score_raw, nca_risk_score_raw, 
                                      pvf_risk_score_raw, 
                                      ## risk score factors
                                      RA2_Current_Charge_Violent,
                                      RA2_Violent_and_under_21, 
                                      RA2_Age_At_Interview,
                                      RA2_New_Charge_While_Case_Pending,
                                      RA2_Has_Prior_Misdemeanor, 
                                      RA2_Has_Prior_Felony,
                                      RA2_How_Many_FTA_Last_2_Yrs,
                                      RA2_Has_Any_FTA_Older_2_Yrs,
                                      RA2_Num_Prior_Violent,
                                      RA2_Has_Prior_Incarceration) %>% 
  mutate(FTA = ifelse(Fta == "Y", 1, 0),
         probation = ifelse(OnProbation == "Y"|SupProbation == "Y", 1, 0),
         Interview_Date = as.Date(Interview_Date_Created),
         PRIM_Case_Filing_Date = as.Date(Case_Date_Booked),
         Age_at_PRIM_Case_Filing = Age_at_Case_Booked,
         fta_risk_score_raw = ifelse(fta_risk_score_raw == "NULL", NA, fta_risk_score_raw),
         nca_risk_score_raw = ifelse(nca_risk_score_raw == "NULL", NA, nca_risk_score_raw),
         pvf_risk_score_raw = ifelse(pvf_risk_score_raw == "NULL", NA, pvf_risk_score_raw)) %>%
  filter(Interview_Date >= "2013-07-01") %>% 
  select(-c(Fta,OnProbation, SupProbation, Case_Date_Booked, 
            Age_at_Case_Booked, Interview_Date_Created))
```

```{r}
summary(PRIM_features0)
```


```{r}
PRIM_features0 %>% group_by(PersonID, SeqCaseID) %>% summarise(count = n ()) %>% arrange(desc(count))
```

__NOTE__:
One case in CourtNet should be matched exactlt to one case in PRIM, using SeqCaseID. The case booked date in PRIM system should be slightly later than the case booked date in the CourtNet system. However, there are multiple reasons why there are many caseIDs corresponding to one SeqCaseID in the PRIM system:

1. for the original case, the defandent may fail to appear. Then the defendant will be arrested again and a new interview and a new case will be created. The new CaseID in the PRIM system is still under the same SeqCaseID from the original case. 
   - PersonID == 5; SeqCaseID == 3046161

2. There are cases when multiple CaseID's are related to one SeqCaseID, which means that there should be more than one pretrial interview related to that one SeqCaseID based on the explanation. However, the interviewID is the same for all CaseID's. Anything wrong here? Is that because there are multiple charges? For rearrest?
   - PersonID == 409241; SeqCaseID == 1719208
   - PersonID == 152948; SeqCaseID == 913837

The reason why we saw different risk scores for the child CaseID's of the same SeqCaseID is that if the FTA appears in the previous case, which is one the factors to calculate risk score, the risk score will definitely change. 

3. PRIM_Cases data set has the same age issue, that is the age at PRIM case filing ranges from 0 to 237, which is impossible. Need to compare the two data sets to see which one is relatively more accurate


###### PRIM Risk Score #######################################################################
## Calculate risk scores

- How_Many_FTA_Last_2_Yrs: from arnold psa, the judging criteria are: 0, 1, 2 or more. From the data, we have: 0, 1 or 2, 3 or more. Conflict. I treat "1 or 2" choice in the data set as the "1" choice from arnold.

## FTA: failure to appear
```{r}
FTA_table = PRIM_features0 %>% 
  mutate(RA2_Has_Prior_Conviction = 
             ifelse(RA2_Has_Prior_Misdemeanor=="2" & RA2_Has_Prior_Felony == "2", 0,   ## "2": No
             ifelse(RA2_Has_Prior_Misdemeanor == "1" |  RA2_Has_Prior_Felony=="1", 1, NA)),
         
         RA2_New_Charge_While_Case_Pending = ifelse(RA2_New_Charge_While_Case_Pending == "1", 1, 
                                             ifelse(RA2_New_Charge_While_Case_Pending == "2", 0, NA)),
         
         RA2_How_Many_FTA_Last_2_Yrs = ifelse(RA2_How_Many_FTA_Last_2_Yrs == "1", 0,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "2", 2,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "3", 4, NA))),
         
         RA2_Has_Any_FTA_Older_2_Yrs = ifelse(RA2_Has_Any_FTA_Older_2_Yrs == "1", 1,
                                       ifelse(RA2_Has_Any_FTA_Older_2_Yrs == "2", 0, NA))) %>%
  select(RA2_New_Charge_While_Case_Pending, 
         RA2_Has_Prior_Conviction, 
         RA2_How_Many_FTA_Last_2_Yrs,
         RA2_Has_Any_FTA_Older_2_Yrs)

### Calculate Scores
FTA_table$Score = apply(FTA_table, 1, sum)
```


## NCA: new criminal activity
```{r}
### Pick NCA Features
NCA_table = PRIM_features0 %>% 
  mutate(RA2_Age_At_Interview = ifelse(RA2_Age_At_Interview == "3", 0,
                                ifelse(RA2_Age_At_Interview == "NULL", NA, 2)),
         
         RA2_New_Charge_While_Case_Pending = ifelse(RA2_New_Charge_While_Case_Pending == "1", 3, 
                                             ifelse(RA2_New_Charge_While_Case_Pending == "2", 0, NA)),
         
         RA2_Has_Prior_Misdemeanor = ifelse(RA2_Has_Prior_Misdemeanor == "1", 1,
                                     ifelse(RA2_Has_Prior_Misdemeanor == "2", 0, NA)),
         
         RA2_Has_Prior_Felony = ifelse(RA2_Has_Prior_Felony == "1", 1,
                                ifelse(RA2_Has_Prior_Felony == "2", 0, NA)),
         
         RA2_Num_Prior_Violent = ifelse(RA2_Num_Prior_Violent == "1", 0,
                                 ifelse(RA2_Num_Prior_Violent == "2", 1,
                                 ifelse(RA2_Num_Prior_Violent == "3", 2, NA))),
         
         RA2_How_Many_FTA_Last_2_Yrs = ifelse(RA2_How_Many_FTA_Last_2_Yrs == "1", 0,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "2", 1,
                                       ifelse(RA2_How_Many_FTA_Last_2_Yrs == "3", 2, NA))),
         
         RA2_Has_Prior_Incarceration = ifelse(RA2_Has_Prior_Incarceration == "1", 2,
                                       ifelse(RA2_Has_Prior_Incarceration == "2", 0, NA))) %>%
  select(RA2_Age_At_Interview, RA2_New_Charge_While_Case_Pending, RA2_Has_Prior_Felony,
         RA2_Has_Prior_Misdemeanor, RA2_Num_Prior_Violent, RA2_How_Many_FTA_Last_2_Yrs,
         RA2_Has_Prior_Incarceration)

## Calculate unweighted Score
NCA_table$Score = apply(NCA_table, 1, sum)
```


## PVF (NVCA)
```{r}
### pick up NVCA features
PVF_table = PRIM_features0 %>% 
  mutate(RA2_Current_Charge_Violent = ifelse(RA2_Current_Charge_Violent == "2", 0, 
                                      ifelse(RA2_Current_Charge_Violent == "1", 2, NA)),
         
         RA2_Violent_and_under_21 = ifelse(RA2_Violent_and_under_21 == "2", 0,
                                    ifelse(RA2_Violent_and_under_21 == "1", 1, NA)),
         
         RA2_New_Charge_While_Case_Pending = ifelse(RA2_New_Charge_While_Case_Pending == "2", 0,
                                             ifelse(RA2_New_Charge_While_Case_Pending == "1",1, NA)),
         
         RA2_Has_Prior_Conviction = 
             ifelse(RA2_Has_Prior_Misdemeanor=="2" & RA2_Has_Prior_Felony == "2", 0, 
             ifelse(RA2_Has_Prior_Misdemeanor == "1" | RA2_Has_Prior_Felony=="1", 1, NA)),
         
         RA2_Num_Prior_Violent = ifelse(RA2_Num_Prior_Violent == "1", 0,
                                 ifelse(RA2_Num_Prior_Violent == "2", 1,
                                 ifelse(RA2_Num_Prior_Violent == "3", 2, NA)))) %>% 
  select(RA2_Current_Charge_Violent, 
         RA2_Violent_and_under_21,
         RA2_New_Charge_While_Case_Pending,
         RA2_Has_Prior_Conviction,
         RA2_Num_Prior_Violent)

## NVCA unweighted score
PVF_table$Score = apply(PVF_table, 1, sum)
```


#### add three calculated scores together 

In this step, risk scores that are NA are removed. In order to test Arnold PSA, we need to have those values. And in order to be consistent with the data sets that we use to train Arnold PSA and all other models, we need to get rid of the NAs as well.


```{r}
PRIM_features = PRIM_features0 %>% 
  select(PersonID, SeqCaseID, CaseID, 
         Interview_Date,
         FTA, probation, 
         RA2_New_Charge_While_Case_Pending,
         RA2_Current_Charge_Violent,
         PRIM_Case_Filing_Date, 
         Age_at_PRIM_Case_Filing,
         fta_risk_score_raw, 
         nca_risk_score_raw, 
         pvf_risk_score_raw) %>% 
  mutate(pending_charge = ifelse(RA2_New_Charge_While_Case_Pending == "2", 0, 1),
         is_violence_prim = ifelse(RA2_Current_Charge_Violent == "2", 0, 1),
         fta_risk_score_raw = as.numeric(fta_risk_score_raw),
         nca_risk_score_raw = as.numeric(nca_risk_score_raw),
         pvf_risk_score_raw = as.numeric(pvf_risk_score_raw),
         fta_calc = FTA_table$Score,
         nca_calc = NCA_table$Score,
         pvf_calc = PVF_table$Score) %>% 
  select(-RA2_New_Charge_While_Case_Pending, -RA2_Current_Charge_Violent) %>%
  na.omit(.)
```


```{r}
summary(PRIM_features)
```



## check age distribution again
```{r}
age = PRIM_features %>% 
  filter(Age_at_PRIM_Case_Filing >= 18 & Age_at_PRIM_Case_Filing <= 70) %>% 
  select(Age_at_PRIM_Case_Filing)

age_count = age %>% group_by(Age_at_PRIM_Case_Filing) %>% summarise(count = n())
ggplot(data = age_count, aes(x=Age_at_PRIM_Case_Filing, y=count)) + geom_bar(stat = "identity")
```

__NOTE__:

Previously, in order to be consistent with the population tested on ML models and Arnold PSA, we excluded records without risk scores, which are a large amount of records (1100000). This leads to the result that many young people are exlcuded from our population, making the age distribution different from what we expected. 


###### Combine PRIM_CASES + CNet_features ############################################################

Merge CourNet features and PRIM features. Get rid of null charge date and change the formate to date format.

```{r}
all_features0 = merge(x=CNet_features, y=PRIM_features, by=c("PersonID", "SeqCaseID"))
```


## check violence feature
```{r}
violence_check = all_features0 %>% 
  select(PersonID, SeqCaseID, ChargeEntityNumber, ChargeDate,is_violence, is_violence_prim)

mean(violence_check$is_violence)
mean(violence_check$is_violence_prim)
```

__NOTE__:

By checking the violence feature from CourtNet and from PRIM, PRIM has more records of 1 when CNet has value of 0. It makes sense since when checking if a charge is violent or not in CourtNet system, we detect a certain string in UOR Description. But PRIM directly has the answer from interview. Therefore, I decided to proceed with PRIM feature.

Additionally, there is a huge difference between the two features. The proportion from CNet feature only has 2.4%, while PRIM has 18.6%. 

```{r}
all_features0 = all_features0 %>% select(-is_violence) %>% rename(is_violence = is_violence_prim)
```


## sanity check
```{r}
a = all_features0[all_features0$PersonID == "5" & all_features0$SeqCaseID == "3046161", ]
b = PRIM_features[PRIM_features$PersonID == "5"& PRIM_features$SeqCaseID == "3046161",]
c = CNet[CNet$PersonID == "5" & CNet$SeqCaseID == "3046161", ]
```


__NOTE__:
There are extra records added to the data set. It is because one SeqCaseID could relate to multiple CaseID's in the PRIM system. Additionally, one SeqCaseID could have multiple charges. Therefore, each one of the CaseIDs will have the same number of charges information added.

Take the example above for instance, the individual with PersonID == 5 & SeqCaseID == 3046161 had 5 charges under this CourtNet case. This CourtNet case (SeqCaseID) had 2 CaseIDs corresponded in PRIM system. Therefore, after inner joining them, there are 10 records in total, since each CaseID will have 5 charges right now. The 5 charges are exactly the same in both CaseIDs. `This is very important since it will change the way we calcualted features later on.`



##### age issue 
```{r}
age_issue = all_features0 %>% select(PersonID, SeqCaseID, CaseID, 
                            CNet_Case_Filing_Date, Age_at_CNet_Case_Filing, 
                            PRIM_Case_Filing_Date,Age_at_PRIM_Case_Filing, ChargeDate) 
sum(age_issue$Age_at_CNet_Case_Filing < 18 | age_issue$Age_at_CNet_Case_Filing > 70)
sum(age_issue$Age_at_PRIM_Case_Filing < 18 | age_issue$Age_at_PRIM_Case_Filing > 70)
```



Checked the two sides of both "age_at_CNet_case_filing" and "age_at_PRIM_case_filing". There are more values of CNet age lying within the normal range. 

Therefore, to get each individual's DOB and age at each charge date I will still proceed with CNet_Case_Filing_Date and Age_at_CNet_Case_Filing Date. Temporarily, I will keep the age_at_current_charge between 18 and 70. If we filter base on PRIM_Case_Filing_Age and then calculate the age_at_charge, there are values ranging from -8000 to 10820, due to the reason that some charges dates could date back to a long time ago.

Additionally, since different PRIM_case_filing_date could have the same filing_age, there might be different DOBs for one person. (PersonID = 8, SeqCaseID = 2523727)


```{r}
all_features = all_features0 %>% 
    mutate(DOB = CNet_Case_Filing_Date - years(Age_at_CNet_Case_Filing),
           age_at_charge = floor(as.numeric(as.period(interval(DOB, ChargeDate)), "years"))) %>%
    filter (age_at_charge >= 18,
            age_at_charge <= 70) %>% 
    select(-c(CNet_Case_Filing_Date, 
              Age_at_CNet_Case_Filing,
              Age_at_PRIM_Case_Filing, DOB,
              Interview_Date))

## re-order the columns 
all_features = all_features[, c(1,2,30,3:5,42,33,6:29,31:32,37:38,34:36,39:41)]
```

```{r}
summary(all_features)
```

## check age_at_charge distribution
```{r}
age_count = all_features %>% group_by(age_at_charge) %>% summarise(count = n())
ggplot(data = age_count, aes(x=age_at_charge, y=count)) + geom_bar(stat = "identity")
```







###### Find cutoff date and screening date

- Cutoff Date: two years eariler than the latest charge date in our data set

```{r}
## cutoff date
cutoff_date = max(all_features$ChargeDate) - years(2)

## subset data
before_cutoff0 = all_features %>% filter(ChargeDate < cutoff_date)
after_cutoff0 = all_features %>% filter(ChargeDate > cutoff_date)
```


```{r}
## cutoff date
#max_date = max(all_features$ChargeDate) - years(1)
#min_cutoff_date = max(all_features$ChargeDate) - years(3)

## subset data
#before_cutoff0 = all_features %>% filter(ChargeDate < min_cutoff_date)
#after_cutoff0 = all_features %>% filter(ChargeDate > min_cutoff_date & ChargeDate < max_date)
```



- Screening Dates: for each individual, the cloesest charge date to the cutoff date will be used as the "screening date". Before and on the screening date, the information will be used as criminal history. After the screening date, the information will be used to determine recidivism.


```{r}
## screening dates
screening_dates = before_cutoff0 %>% 
    group_by(PersonID) %>% 
    summarise(screening_date = max(ChargeDate))

## merge screening dates with before_cutoff data set
before_cutoff0 = merge(x=before_cutoff0, y=screening_dates, by="PersonID", all.x = T) %>% 
    mutate(Gender = ifelse(Gender == "Male", 1, 0))
```

#### Add year indicator features

Since we have charge dates and screening dates, now we can calculate the time interval between each charge date and the screening date and then determine whether or not the person had committed crime in the past 5 years, 3 years, 1 year, and 6 months.

The year_case variable is for each PRIM case, since the FTA and probation information is based on PRIM case level. Using the year_case, we can determine the other two features: "fta_two_year" and "fta_two_year_plus".

```{r}
before_cutoff = before_cutoff0[, c(1,43,2:42)] %>% 
    mutate(year_charges = as.numeric(as.period(interval(ChargeDate, screening_date)), "years"),
           year_cases = as.numeric(as.period(interval(PRIM_Case_Filing_Date, 
                                                      screening_date)), "years"),
           fta_two_year = ifelse(FTA == 1 & year_cases <= 2, 1, 0),
           fta_two_year_plus = ifelse(FTA == 1 & year_cases > 2, 1, 0),
           six_month = ifelse(year_charges <= 0.5, 1, 0),
           one_year = ifelse(year_charges <= 1, 1, 0),
           three_year = ifelse(year_charges <= 3, 1, 0),
           five_year = ifelse(year_charges <= 5, 1, 0)) %>% 
    select(-c(FTA, year_charges, year_cases, PRIM_Case_Filing_Date))
```

```{r}
summary(before_cutoff)
```







### function to create features
```{r}
create_features = function(data){

## first layer of group: for fta & probation information, since they are based on CaseID level. 
    
    features = data %>% 
        group_by(PersonID, screening_date, SeqCaseID, CaseID, Gender) %>% 
        summarise(age_at_current_charge = max(age_at_charge),
                  p_SentMonths = max(SentMonths),
                  p_charges = max(ChargeEntityNumber) - min(ChargeEntityNumber)+1,
                  p_violence = sum(is_violence),
                  p_felony = sum(is_felony),
                  p_misdemeanor = sum(is_misdemeanor),
                  p_property = sum(is_property),
                  p_murder = sum(is_murder),
                  p_assault = sum(is_assault),
                  p_sex_offense = sum(is_sex_offense),
                  p_weapon = sum(is_weapon),
                  p_felprop_viol = sum(is_fel_prop_viol),
                  p_felassult = sum(is_fel_assult),
                  p_misdeassult = sum(is_misde_assult),
                  p_traffic = sum(is_traffic),
                  p_drug = sum(is_drug),
                  p_dui = sum(is_dui),
                  p_stalking = sum(is_stalking),
                  p_voyeurism = sum(is_voyeurism),
                  p_fraud = sum(is_fraud),
                  p_stealing = sum(is_stealing),
                  p_trespass = sum(is_trespass),
                  ADE = max(ADE),
                  Treatment = max(Treatment),
                  p_prison = max(prison), ## even though five charges under one case could have 
                  p_jail = max(jail),     ## three prison/jail sentence, for instance, the person
                                          ## will only have one jail/prison count for that case.
                  
                  ## PRIM Case Level Information
                  p_fta_two_year = max(fta_two_year),
                  p_fta_two_year_plus = max(fta_two_year_plus),
                  p_probation = max(probation),
                  p_pending_charge = max(pending_charge),
                  six_month = max(six_month),     ## as long as there is one charge that is
                  one_year = max(one_year),       ## within six month/one year/three years/five
                  three_year = max(three_year),   ## years, the indicator variable is one.
                  five_year = max(five_year),
                  fta_risk_score_raw = mean(fta_risk_score_raw),
                  nca_risk_score_raw = mean(nca_risk_score_raw),
                  pvf_risk_score_raw = mean(pvf_risk_score_raw),
                  fta_calc = mean(fta_calc),
                  nca_calc = mean(nca_calc),
                  pvf_calc = mean(pvf_calc)) %>% 
## second layer group: for criminal history information, since they are based on SeqCaseID level.
        group_by(PersonID, screening_date, SeqCaseID, Gender) %>% 
        summarise(p_arrest = length(CaseID),
                  p_SentMonths = max(p_SentMonths),
                  age_at_current_charge = max(age_at_current_charge),
                  p_charges = max(p_charges),
                  p_violence = max(p_violence),
                  p_felony = max(p_felony),
                  p_misdemeanor = max(p_misdemeanor),
                  p_property = max(p_property),
                  p_murder = max(p_murder),
                  p_assault = max(p_assault),
                  p_sex_offense = max(p_sex_offense),
                  p_weapon = max(p_weapon),
                  p_felprop_viol = max(p_felprop_viol),
                  p_felassult = max(p_felassult),
                  p_misdeassult = max(p_misdeassult),
                  p_traffic = max(p_traffic),
                  p_drug = max(p_drug),
                  p_dui = max(p_dui),
                  p_stalking = max(p_stalking),
                  p_voyeurism = max(p_voyeurism),
                  p_fraud = max(p_fraud),
                  p_stealing = max(p_stealing),
                  p_trespass = max(p_trespass),
                  ADE = max(ADE),
                  Treatment = max(Treatment),
                  p_prison = max(p_prison),
                  p_jail = max(p_jail),
                  
                  ## CourtNet Case Level Information
                  p_fta_two_year = sum(p_fta_two_year),
                  p_fta_two_year_plus = sum(p_fta_two_year_plus),
                  p_probation = sum(p_probation),
                  p_pending_charge = sum(p_pending_charge),
                  six_month = max(six_month),
                  one_year = max(one_year),
                  three_year = max(three_year),
                  five_year = max(five_year),
                  fta_risk_score_raw = max(fta_risk_score_raw),
                  nca_risk_score_raw = max(nca_risk_score_raw),
                  pvf_risk_score_raw = max(pvf_risk_score_raw),
                  fta_calc = max(fta_calc),
                  nca_calc = max(nca_calc),
                  pvf_calc = max(pvf_calc)) %>% 
## third layer: individual level; sum up all the SeqCaseID information related to one person.
        group_by(PersonID, Gender, screening_date) %>%   
        summarise(age_at_current_charge = max(age_at_current_charge),
                  p_SentMonths = sum(p_SentMonths),
                  p_arrest = sum(p_arrest),
                  p_charges = sum(p_charges),
                  p_violence = sum(p_violence),
                  p_felony = sum(p_felony),
                  p_misdemeanor = sum(p_misdemeanor),
                  p_property = sum(p_property),
                  p_murder = sum(p_murder),
                  p_assault = sum(p_assault),
                  p_sex_offense = sum(p_sex_offense),
                  p_weapon = sum(p_weapon),
                  p_felprop_viol = sum(p_felprop_viol),
                  p_felassult = sum(p_felassult),
                  p_misdeassult = sum(p_misdeassult),
                  p_traffic = sum(p_traffic),
                  p_drug = sum(p_drug),
                  p_dui = sum(p_dui),
                  p_stalking = sum(p_stalking),
                  p_voyeurism = sum(p_voyeurism),
                  p_fraud = sum(p_fraud),
                  p_stealing = sum(p_stealing),
                  p_trespass = sum(p_trespass),
                  ADE = sum(ADE),  
                  Treatment = sum(Treatment),
                  p_prison = sum(p_prison),
                  p_jail = sum(p_jail),
                  p_fta_two_year = sum(p_fta_two_year),
                  p_fta_two_year_plus = sum(p_fta_two_year_plus),
                  p_pending_charge = sum(p_pending_charge),
                  p_probation = sum(p_probation),
                  six_month = max(six_month),
                  one_year = max(one_year),
                  three_year = max(three_year),
                  five_year = max(five_year),
                  fta_risk_score_raw = max(fta_risk_score_raw),
                  nca_risk_score_raw = max(nca_risk_score_raw),
                  pvf_risk_score_raw = max(pvf_risk_score_raw),
                  fta_calc = max(fta_calc),
                  nca_calc = max(nca_calc),
                  pvf_calc = max(pvf_calc))
    return(features)
}
```


#### Before Screening Date ###########################################################################

```{r}
data_before = before_cutoff %>% filter(ChargeDate < screening_date)
features_before = create_features(data_before) %>% 
    select(PersonID, Gender, screening_date, 
           p_SentMonths, six_month, one_year, three_year, five_year)
```


##### On Screening Date ####################################################################

The screening_date can be treated as the "current offense" date and the data before this date will be treated as criminal history information. Notice that since one SeqCaseID can have multiple CaseID's, therefore all these CaseIDs share the same charge information for that one single SeqCaseID. If we sum up the features based on CaseID level, the results are way larger than the actual number. Therefore, we sum up the features based on CaseID level, and divided the number of CaseIDs. 

Three layer group reason: 

- FTA: CaseID level
- probation: CaseID level
- Other features: charge level

For sentence months, even though it is at charge level, we do not use the sum of sentence month for each charge under each SeqCaseID. Instead, we use the max sentence month. Because sentence is not a cummulative activity (like you stay in jail for 48 years for the first charge and then stay for the second one for 24 years). Using the max sentence makes more sense. For other features, such as is_traffic, is_drug etc., they are also based on charge level. But summing those binary values makes sense since each charge under the same case could be different. 

-- tested with: PersonID == 79, SeqCaseID == 425760; CaseID == 1255640

-- For charges on screening date, which means they are current charges, we would like to know if the current charge involves "violence". 


```{r}
data_on = before_cutoff %>% filter(ChargeDate == screening_date)
features_on = data_on %>% 
    group_by(PersonID) %>% 
    summarise(current_violence = max(is_violence))
```



##### Before and On Screening Date ###################################################

The last step was to further improve the data quality. The p_SentMonths is the total amount of jail/prison time that each individual had before the current charge. Since our data set excluded juvnile cases, therefore, we would expect that the age at current offense to be equal to or larger than the privious sentenced years + 18 years. If not, then there must be something wrong with the sentence month information, and the record needs to be removed.


```{r}
data_before_on = before_cutoff %>% filter(ChargeDate <= screening_date)
features_before_on = create_features(data_before_on) %>% 
    select(-c(six_month, one_year, three_year, five_year, p_SentMonths)) %>% 
    merge(x = ., y = features_before, by = "PersonID", all.x = TRUE) %>% 
    ## if all the following variables are NA, it means that those individuals do not have any
    ## criminal history besides the current charges.
    mutate(six_month = ifelse(is.na(six_month), 0, six_month),
           one_year = ifelse(is.na(one_year), 0, one_year),
           three_year = ifelse(is.na(three_year), 0, three_year),
           five_year = ifelse(is.na(five_year), 0, five_year),
           p_SentMonths = ifelse(is.na(p_SentMonths), 0, p_SentMonths)) %>% 
    select(-c(Gender.y, screening_date.y)) %>% 
    rename(Gender = Gender.x,
           screening_date = screening_date.x) %>% 
    merge(x=., y=features_on, by = "PersonID", all.x = TRUE) %>% 
    filter(age_at_current_charge >= p_SentMonths/12 + 18)

features_before_on = features_before_on[,c(1:33,40:45,34:39)]
```

```{r}
summary(features_before_on)
```


## check age distribution
```{r}
age_count = features_before_on %>% group_by(age_at_current_charge) %>% summarise(count = n())
ggplot(data = age_count, aes(x=age_at_current_charge, y=count)) + geom_bar(stat = "identity")
```











#### After Screening Date #################################################################

Since the screening dates are the latest dates of charges for each individual before the cutoff date, therefore, data after screening date is the same as data after the cutoff date. From now on, we still start using "after_cutoff" data set to create different prediction labels.

```{r}
## inner join two data sets
## if there is no PersonID from screening dates that have match in the after_cutoff data set, then those PersonID's did not commit any crime after the cutoff_date, or the screening date.


## sentence month information on the screening date (current offense date). We need this information to determine the date that the person was released and then see if any crime has been commited two years after this date.

release_dates = data_on %>% 
    group_by(PersonID, screening_date) %>% 
    summarise(current_SentMonth = max(SentMonths),
              current_SuspSentMonth = max(SuspSentMonths)) %>% 
    mutate(SentDays = ifelse(current_SentMonth>0 & current_SuspSentMonth==0,
                             round(current_SentMonth*30.8333), 0),
           release_date = screening_date+SentDays) %>% 
    select(PersonID, release_date, current_SentMonth, current_SuspSentMonth)

## merge
after_cutoff = merge(x=after_cutoff0, y=release_dates, by="PersonID", all.y = TRUE) %>% 
    filter(current_SentMonth <= 12 | (current_SentMonth > 12 & current_SuspSentMonth > 0)) %>% 
    select(-current_SentMonth, -current_SuspSentMonth)

## add year feature
data_after = after_cutoff[, c(1,43,2:42)] %>% 
    mutate(year_offenses = as.numeric(as.period(interval(release_date, ChargeDate)), "years"),
           recid_two_year = ifelse(year_offenses <= 2, 1, 0),
           recid_six_month = ifelse(year_offenses <= 0.5, 1, 0),
           recid_drug = ifelse(recid_two_year == 1 & is_drug == 1, 1, 0),
           recid_traffic = ifelse(recid_two_year == 1 & is_traffic == 1, 1, 0),
           recid_violence = ifelse(recid_two_year == 1 & is_violence == 1, 1, 0),
           recid_F = ifelse(recid_two_year == 1 & is_felony == 1, 1, 0),
           recid_M = ifelse(recid_two_year == 1 & is_misdemeanor == 1, 1, 0),
           recid_stealing = ifelse(recid_two_year == 1 & is_stealing == 1, 1, 0),
           recid_property = ifelse(recid_two_year == 1 & is_property == 1, 1, 0)) %>% 
    select(PersonID:ChargeDate, ChargeEntityNumber, year_offenses:recid_property)

```



###### Generate labels ########################################################################

Using max() function because as long as there is one charge that is within two years of the screening date, the person has recidivism. 

```{r}
outcomes = data_after %>% 
    group_by(PersonID) %>% 
    summarise(recid_two_year = max(recid_two_year), 
              recid_six_month = max(recid_six_month),
              recid_drug = max(recid_drug),
              recid_traffic = max(recid_traffic),
              recid_violence = max(recid_violence),
              recid_F = max(recid_F),
              recid_M = max(recid_M),
              recid_stealing = max(recid_stealing),
              recid_property = max(recid_property)) 

```

#### Combine criminal history data with labels 

```{r}
recid_labels = merge(x=features_before_on, y=outcomes, by = "PersonID") %>% 
    mutate(recid_two_year = ifelse(is.na(recid_two_year), 0,  recid_two_year),
           recid_six_month = ifelse(is.na(recid_six_month), 0,  recid_six_month),
           recid_drug = ifelse(is.na(recid_drug), 0,  recid_drug),
           recid_traffic = ifelse(is.na(recid_traffic), 0,  recid_traffic),
           recid_violence = ifelse(is.na(recid_violence), 0,  recid_violence),
           recid_F = ifelse(is.na(recid_F), 0,  recid_F),
           recid_M = ifelse(is.na(recid_M), 0,  recid_M),
           recid_stealing = ifelse(is.na(recid_stealing), 0,  recid_stealing),
           recid_property = ifelse(is.na(recid_property), 0,  recid_property))
```

```{r}
summary(recid_labels)
```

```{r}
mean(recid_labels$recid_two_year == 1)
mean(recid_labels$recid_six_month == 1)
```


```{r}
age_dist = recid_labels %>% 
  select(age_at_current_charge, recid_two_year, recid_six_month) %>% 
  group_by(age_at_current_charge) %>% 
  summarise(recid_two_year_prob = mean(recid_two_year == 1))

ggplot(data = age_dist, aes(x=age_at_current_charge, y = recid_two_year_prob)) +
  geom_bar(stat = "identity")
```

## visualization 1

Population: individuals who have one or more drug-related crimes or people who have ADE

```{r}
pop1 = recid_labels %>% filter(p_drug >= 1, ADE >= 1)
age_dist1 = pop1 %>% 
  select(age_at_current_charge, recid_two_year) %>% 
  group_by(age_at_current_charge) %>% 
  summarise(recid_two_year_prob = mean(recid_two_year == 1))
ggplot(data = age_dist1, aes(x=age_at_current_charge, y = recid_two_year_prob)) +
  geom_bar(stat = "identity")
```

## visualization 2

Population: individuals who don't have drug-related crimes or ADE

```{r}
pop2 = recid_labels %>% filter(p_drug == 0 & ADE == 0)
age_dist2 = pop2 %>% 
  select(age_at_current_charge, recid_two_year, recid_six_month) %>% 
  group_by(age_at_current_charge) %>% 
  summarise(recid_two_year_prob = mean(recid_two_year == 1))

ggplot() +
  geom_bar(data = age_dist2, 
           aes(x=age_at_current_charge, y = recid_two_year_prob), 
           stat = "identity")

ggplot() +
  geom_bar(data = age_dist2, 
           aes(x=age_at_current_charge, y = recid_two_year_prob, fill="red"), 
           stat = "identity") +
  geom_bar(data = age_dist1, 
           aes(x=age_at_current_charge, y = recid_two_year_prob, fill = "blue"), 
           stat = "identity")
```


#### Save data files
```{r}
save(data_before,features_before, 
     data_on, features_on,
     data_before_on, data_after, features_before_on, 
     outcomes, recid_labels,
     file = "kentucy_features.Rdata")
```

```{r}
## in order to replicate the spliting process for train/test data sets, set seed = 816. 

set.seed(816)
train_index = sample(1:nrow(recid_labels), 0.9*nrow(recid_labels), replace = F)
train = recid_labels[train_index,]
test = recid_labels[-train_index,]

write.csv(recid_labels, file = "C:/Users/binha/Documents/Duke/Cynthia Research/KY-analysis-mytrials/KY Recidivism/KY data/kentucky_data.csv", row.names = F)
write.csv(train, file = "C:/Users/binha/Documents/Duke/Cynthia Research/KY-analysis-mytrials/KY Recidivism/KY data/train.csv", row.names = F)
write.csv(test, file = "C:/Users/binha/Documents/Duke/Cynthia Research/KY-analysis-mytrials/KY Recidivism/KY data/test.csv", row.names = F)
```














